### 2024-09-13 project/proposal.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}

\title{Proposal Submission for Deep Learning Project}
\author{
    Group Members: \\
    Farooq Mahmud
}
\date{}

\begin{document}

\maketitle

\section{Introduction}
For my deep learning project, I will analyze and replicate the seminal deep learning paper, \textit{"Going Deeper with Convolutions"} by Christian Szegedy et al., presented at the CVPR 2015 conference. This paper introduces the Inception architecture, which revolutionized deep learning by addressing the challenge of building efficient and scalable neural networks.

The primary aim of this project is to implement the \textit{GoogLeNet} architecture described in the paper using TensorFlow, replicate its performance on the ImageNet dataset, and evaluate its generalization capability by applying it to the \textit{CIFAR-10} dataset. Additionally, I will provide a critical analysis of the impact of the Inception architecture on the field of deep learning, highlighting its influence on subsequent research and applications.

\section{Objectives}
The key objectives of our project are as follows:
\begin{itemize}
    \item \textbf{Understand and articulate} the core ideas and contributions of the paper, including the use of Inception modules for building efficient and scalable neural networks.
    \item \textbf{Implement the GoogLeNet architecture} using TensorFlow and train the model on the ImageNet dataset (or a subset thereof).
    \item \textbf{Replicate the results} reported in the paper, including the model's classification performance on ImageNet.
    \item \textbf{Generalize the model} by applying it to the CIFAR-10 dataset and evaluate its performance on this smaller dataset.
    \item \textbf{Analyze the paper's impact} on deep learning, particularly in computer vision, by reviewing citations and subsequent research that built on this work.
    \item \textbf{Document the challenges} encountered during implementation, training, and testing.
\end{itemize}

\end{document}

### 2024-09-15 project/utils/download_synsets.py
Add async script to download and rename synset tar files
Introduced a new script `download_synsets.py` to automate the download and renaming of ImageNet synset tar files using `aiohttp` and `asyncio`. The script includes error handling, logging, and creates required directories if they do not exist.

@@ -1,0 +1,71 @@
+import os
+import aiohttp
+import asyncio
+import logging
+
+logging.basicConfig(level=logging.INFO,
+                    format="%(asctime)s - %(levelname)s - %(message)s")
+
+ROOT_URL = "https://image-net.org/data/winter21_whole/"
+
+synset_mapping = {
+    "airplane": "n02691156",
+    "automobile": "n02958343",
+    "bird": "n01503061",
+    "cat": "n02121808",
+    "deer": "n02419796",
+    "dog": "n02084071",
+    "frog": "n01641577",
+    "horse": "n02374451",
+    "ship": "n04194289",
+    "truck": "n04467665"
+}
+
+download_dir = r"C:\Users\faroo\Downloads\imagenet_cifar10_tars"
+
+if not os.path.exists(download_dir):
+    os.makedirs(download_dir)
+
+
+async def download_and_rename_tar(session, synset_id, cifar_class_name):
+    tar_url = f"{ROOT_URL}{synset_id}.tar"
+    renamed_tar_filename = f"{synset_id}-{cifar_class_name}.tar"
+    renamed_tar_file_path = os.path.join(download_dir, renamed_tar_filename)
+
+    if os.path.exists(renamed_tar_file_path):
+        logging.info(f"{renamed_tar_filename} already exists, skipping download.")
+        return
+
+    try:
+        async with session.get(tar_url) as response:
+            if response.status == 200:
+                logging.info(f"Downloading {tar_url}...")
+                tar_file_path = os.path.join(download_dir, f"{synset_id}.tar")
+                with open(tar_file_path, 'wb') as f:
+                    while True:
+                        chunk = await response.content.read(1024)
+                        if not chunk:
+                            break
+                        f.write(chunk)
+
+                logging.info(f"Downloaded {tar_file_path}")
+                os.rename(tar_file_path, renamed_tar_file_path)
+                logging.info(f"Renamed {tar_file_path} to {renamed_tar_filename}")
+            else:
+                logging.error(
+                    f"Failed to download {tar_url}, status code: {response.status}")
+    except Exception as e:
+        logging.error(f"An error occurred while downloading {tar_url}: {e}")
+
+
+async def main():
+    async with aiohttp.ClientSession() as session:
+        tasks = []
+        for cifar_class, synset_id in synset_mapping.items():
+            tasks.append(download_and_rename_tar(session, synset_id, cifar_class))
+
[…]

### 2024-09-15 project/utils/download_synsets.py
Add pyproject.toml and enhance download_synsets.py docs.
Added pyproject.toml with ruff linting settings. Enhanced documentation in download_synsets.py to improve code clarity and usability, and added missing comma in synset_mapping dictionary.

@@ -1,7 +1,46 @@
-import os
-import aiohttp
+"""
+Module: download_synsets
+
+This module provides functionality to download and rename synset tar files from
+ImageNet.
+It utilizes asynchronous operations to handle multiple downloads concurrently, making
+the download process efficient and quicker.
+
+Logging is configured to display info level messages which will help in tracking the
+progress of downloads.
+
+The module defines a main function which orchestrates the download process for all the
+synsets specified in the synset_mapping dictionary. Each synset is downloaded and
+renamed according to its CIFAR-10 class name.
+
+The primary components of this module include:
+
+- ROOT_URL: The base URL from where the ImageNet tar files are downloaded.
+- synset_mapping: A dictionary mapping CIFAR-10 class names to their corresponding
+  synset identifiers in ImageNet.
+- download_dir: The directory where the downloaded tar files will be stored.
+
+Functions:
+- download_and_rename_tar: This function handles the actual downloading and renaming
+  of a synset tar file using aiohttp for asynchronous HTTP requests.
+- main: The main function that initializes the aiohttp ClientSession and triggers
+  the download of all synsets concurrently.
+
+Usage:
+Run the module as a standalone script to initiate the download process:
+    $ python download_synsets.py
+
+Dependencies:
+- aiohttp: Asynchronous HTTP client/server framework
+- asyncio: Python standard library module for writing asynchronous code
+- logging: Python standard library module for logging messages
+- os: Python standard library module for operating system-related functionalities
+"""
 import asyncio
 import logging
+import os
+
+import aiohttp
 
 logging.basicConfig(level=logging.INFO,
                     format="%(asctime)s - %(levelname)s - %(message)s")
@@ -18,7 +57,7 @@ synset_mapping = {
     "frog": "n01641577",
     "horse": "n02374451",
     "ship": "n04194289",
-    "truck": "n04467665"
+    "truck": "n04467665",
 }
 
 download_dir = r"C:\Users\faroo\Downloads\imagenet_cifar10_tars"
@@ -28,6 +67,16 @@ if not os.path.exists(download_dir):
 
 
 async def download_and_rename_tar(session, synset_id, cifar_class_name):
+    """
+    Download the tar file for a given synset and rename it with the CIFAR-10
+    class name.
+    @param session: An active aiohttp ClientSession used to make HTTP requests.
+    @param synset_id: The identifier of the synset used to create the download URL
+    and the initial tar file name.
[…]

### 2024-09-16 project/utils/download_synsets.py
Add argparse for dynamic download directory
Introduce argparse to allow users to specify the download directory via command-line arguments. This enhances the flexibility of the script by removing the hardcoded download directory and replacing it with a user-defined path.

@@ -36,6 +36,7 @@ Dependencies:
 - logging: Python standard library module for logging messages
 - os: Python standard library module for operating system-related functionalities
 """
+import argparse
 import asyncio
 import logging
 import os
@@ -60,13 +61,7 @@ synset_mapping = {
     "truck": "n04467665",
 }
 
-download_dir = r"C:\Users\faroo\Downloads\imagenet_cifar10_tars"
-
-if not os.path.exists(download_dir):
-    os.makedirs(download_dir)
-
-
-async def download_and_rename_tar(session, synset_id, cifar_class_name):
+async def download_and_rename_tar(session, synset_id, cifar_class_name, download_dir):
     """
     Download the tar file for a given synset and rename it with the CIFAR-10
     class name.
@@ -75,6 +70,7 @@ async def download_and_rename_tar(session, synset_id, cifar_class_name):
     and the initial tar file name.
     @param cifar_class_name: The CIFAR class name used to rename the downloaded
     tar file.
+    @param download_dir: The directory where the downloaded tar file will be stored.
     @return: None
     """
     tar_url = f"{ROOT_URL}{synset_id}.tar"
@@ -107,15 +103,30 @@ async def download_and_rename_tar(session, synset_id, cifar_class_name):
         logging.error(f"An error occurred while downloading {tar_url}: {e}")
 
 
-async def main():
-    """Main function to download all synsets asynchronously."""
+async def main(download_dir: str):
+    """
+    Main function to download all synsets asynchronously.
+     @return: None
+    """
+    if not os.path.exists(download_dir):
+        os.makedirs(download_dir)
+
     async with aiohttp.ClientSession() as session:
         tasks = []
         for cifar_class, synset_id in synset_mapping.items():
-            tasks.append(download_and_rename_tar(session, synset_id, cifar_class))
+            tasks.append(download_and_rename_tar(
+                session, synset_id, cifar_class, download_dir))
 
         await asyncio.gather(*tasks)
 
 
 if __name__ == "__main__":
-    asyncio.run(main())
+    parser = argparse.ArgumentParser(
+        description="Download ImageNet synsets and rename them with CIFAR-10 "
+                    "class names.")
+
+    parser.add_argument("--download-dir", type=str, required=True,
+                        help="Directory where tar files will be downloaded.")
+
+    args = parser.parse_args()
+    asyncio.run(main(args.download_dir))

### 2024-09-16 project/utils/preprocess_images.py
Add image preprocessing script
Introduce a new script for resizing images, supporting recursive directory traversal. The script utilizes multiprocessing for efficiency and includes detailed logging for monitoring purposes. Users can specify input/output directories, target size, and number of workers via command-line arguments.

@@ -1,0 +1,98 @@
+import os
+import argparse
+import logging
+import time
+from PIL import Image
+from concurrent.futures import ProcessPoolExecutor
+from pathlib import Path
+from collections import defaultdict
+
+logging.basicConfig(
+    level=logging.INFO,
+    format='%(asctime)s - %(levelname)s - %(message)s',
+    handlers=[
+        logging.FileHandler("image_resizer.log"),
+        logging.StreamHandler()
+    ]
+)
+
+
+# Function to resize an image
+def resize_image(image_path, output_path, size):
+    try:
+        with Image.open(image_path) as img:
+            resized_img = img.resize(size,  Image.Resampling.LANCZOS)
+            output_path.parent.mkdir(parents=True, exist_ok=True)
+            resized_img.save(output_path)
+        logging.info(f"Resized and saved: {output_path}")
+    except Exception as e:
+        logging.error(f"Error resizing image {image_path}: {e}")
+
+
+def find_images(input_dir):
+    image_extensions = ('.png', '.jpg', '.jpeg', '.bmp', '.tiff')
+    image_count_by_type = defaultdict(int)
+    image_paths = []
+
+    for file in Path(input_dir).rglob('*'):
+        if file.suffix.lower() in image_extensions:
+            image_paths.append(file)
+            image_count_by_type[file.suffix.lower()] += 1  # Count each image type
+
+    return image_paths, image_count_by_type
+
+
+def process_images(input_dir, output_dir, size, max_workers):
+    images, image_count_by_type = find_images(input_dir)
+
+    if not images:
+        logging.warning(f"No images found in the directory: {input_dir}")
+        return
+
+    logging.info(
+        f"Found {len(images)} images in {input_dir}. Starting resizing process...")
+
+    for ext, count in image_count_by_type.items():
+        logging.info(f"Found {count} {ext} images.")
+
+    image_tasks = [
+        (image_path, Path(output_dir) / image_path.relative_to(input_dir))
+        for image_path in images
+    ]
+
+    start_time = time.time()
+
+    with ProcessPoolExecutor(max_workers=max_workers) as executor:
+        futures = [executor.submit(resize_image, image_path, output_path, size) for
+                   image_path, output_path in image_tasks]
+
+        for future in futures:
+            future.result()
+
+    end_time = time.time()
+    elapsed_time = end_time - start_time
+    logging.info(
+        f"Time taken to resize {len(images)} images: {elapsed_time:.2f} seconds.")
+
+def main():
+    parser = argparse.ArgumentParser(
[…]

### 2024-09-16 project/utils/preprocess_images.py
Add image preprocessing script
Introduce a new script for resizing images, supporting recursive directory traversal. The script utilizes multiprocessing for efficiency and includes detailed logging for monitoring purposes. Users can specify input/output directories, target size, and number of workers via command-line arguments.

@@ -76,7 +76,7 @@ def process_images(input_dir, output_dir, size, max_workers):
 
 def main():
     parser = argparse.ArgumentParser(
-        description='Resize images in a folder recursively to 244x244.')
+        description='Resize images in a folder recursively.')
     parser.add_argument('--input-dir', type=str, required=True,
                         help='Path to the input folder containing images')
     parser.add_argument('--output-dir', type=str, required=True,

### 2024-09-16 project/utils/preprocess_images.py
Refactor quotes and format code for readability
Standardize quotation marks to double quotes and improve code formatting to enhance readability and consistency throughout the script. This includes adjustments in logging setup, argument parsing, and other string literals.

@@ -1,19 +1,19 @@
-import os
 import argparse
 import logging
 import time
-from PIL import Image
+from collections import defaultdict
 from concurrent.futures import ProcessPoolExecutor
 from pathlib import Path
-from collections import defaultdict
+
+from PIL import Image
 
 logging.basicConfig(
     level=logging.INFO,
-    format='%(asctime)s - %(levelname)s - %(message)s',
+    format="%(asctime)s - %(levelname)s - %(message)s",
     handlers=[
         logging.FileHandler("image_resizer.log"),
-        logging.StreamHandler()
-    ]
+        logging.StreamHandler(),
+    ],
 )
 
 
@@ -30,11 +30,11 @@ def resize_image(image_path, output_path, size):
 
 
 def find_images(input_dir):
-    image_extensions = ('.png', '.jpg', '.jpeg', '.bmp', '.tiff')
+    image_extensions = (".png", ".jpg", ".jpeg", ".bmp", ".tiff")
     image_count_by_type = defaultdict(int)
     image_paths = []
 
-    for file in Path(input_dir).rglob('*'):
+    for file in Path(input_dir).rglob("*"):
         if file.suffix.lower() in image_extensions:
             image_paths.append(file)
             image_count_by_type[file.suffix.lower()] += 1  # Count each image type
@@ -76,23 +76,25 @@ def process_images(input_dir, output_dir, size, max_workers):
 
 def main():
     parser = argparse.ArgumentParser(
-        description='Resize images in a folder recursively.')
-    parser.add_argument('--input-dir', type=str, required=True,
-                        help='Path to the input folder containing images')
-    parser.add_argument('--output-dir', type=str, required=True,
-                        help='Path to the output folder where resized images will be saved')
-    parser.add_argument('--size', type=int, nargs=2, default=(224, 224),
-                        help='Target size for the resized images (width height)')
-    parser.add_argument('--workers', type=int, default=4,
-                        help='Number of parallel workers to use for resizing')
+        description="Resize images in a folder recursively.")
+    parser.add_argument("--input-dir", type=str, required=True,
+                        help="Path to the input folder containing images")
+    parser.add_argument("--output-dir", type=str, required=True,
+                        help="Path to the output folder where resized images "
+                             "will be saved")
+    parser.add_argument("--size", type=int, nargs=2, default=(224, 224),
+                        help="Target size for the resized images (width height)")
[…]

### 2024-09-16 project/utils/preprocess_images.py
Refactor quotes and format code for readability
Standardize quotation marks to double quotes and improve code formatting to enhance readability and consistency throughout the script. This includes adjustments in logging setup, argument parsing, and other string literals.

@@ -17,7 +17,6 @@ logging.basicConfig(
 )
 
 
-# Function to resize an image
 def resize_image(image_path, output_path, size):
     try:
         with Image.open(image_path) as img:

### 2024-09-16 project/utils/preprocess_images.py
Add docstrings to preprocess_images.py functions
Included comprehensive docstrings for functions in preprocess_images.py to enhance code readability and maintainability. These docstrings provide details on the purpose, parameters, and expected behavior of each function.

@@ -1,3 +1,16 @@
+"""
+Module for preprocessing images.
+
+This module includes functions to resize and find images, as well as a main function
+to process images as a whole.
+
+Functions:
+    - resize_image: Resizes an image to given dimensions.
+    - find_images: Identifies and retrieves images from a specified directory.
+    - process_images: Handles the overall processing workflow for images.
+    - main: Entry point for the module, coordinating the complete image
+    preprocessing pipeline.
+"""
 import argparse
 import logging
 import time
@@ -18,6 +31,14 @@ logging.basicConfig(
 
 
 def resize_image(image_path, output_path, size):
+    """
+    Resize an image.
+    @param image_path: The file path of the input image to be resized.
+    @param output_path: The file path where the resized image will be saved.
+    @param size: The dimensions to resize the image to, specified as
+    a tuple (width, height).
+    @return: None
+    """
     try:
         with Image.open(image_path) as img:
             resized_img = img.resize(size,  Image.Resampling.LANCZOS)
@@ -29,6 +50,12 @@ def resize_image(image_path, output_path, size):
 
 
 def find_images(input_dir):
+    """
+    Recursively find images to resize.
+    @param input_dir: Directory to search for image files.
+    @return: A tuple containing a list of image paths and a dictionary
+    with the count of each image type.
+    """
     image_extensions = (".png", ".jpg", ".jpeg", ".bmp", ".tiff")
     image_count_by_type = defaultdict(int)
     image_paths = []
@@ -42,6 +69,14 @@ def find_images(input_dir):
 
 
 def process_images(input_dir, output_dir, size, max_workers):
+    """
+    This method does the work of resizing images.
+    @param input_dir: Directory path where input images are stored
+    @param output_dir: Directory path where resized images will be stored
+    @param size: Desired size (width, height) of the resized images
+    @param max_workers: Maximum number of worker processes to use for resizing
+    @return: None
+    """
     images, image_count_by_type = find_images(input_dir)
 
     if not images:

### 2024-09-18 project/utils/dataset_utils.py
Add train-validation-test split to save_dataset function
Updated save_dataset to split data into training, validation, and test sets. Modified load_dataset to accommodate the additional datasets. Adjusted unit tests to verify the changes and ensure data is correctly split and loaded.

@@ -26,36 +26,48 @@ import argparse
 import pickle
 from pathlib import Path
 from typing import List, Tuple
+from sklearn.model_selection import train_test_split
 
 import tensorflow as tf
 
 
-def save_dataset(input_dir:Path, output_file:Path) -> None:
+def save_dataset(
+        input_dir: Path,
+        output_file: Path,
+        train_size=0.7,
+        val_size=0.2) -> None:
     """
     Save a dataset of images and their corresponding labels to a file.
-
     Args:
         input_dir (Path): The directory containing the image dataset.
         output_file (Path): The file path where the dataset will be saved.
-
+        train_size (float): Proportion of the dataset to include in the training set.
+        val_size (float): Proportion of the dataset to include in the validation set.
     Raises:
-        ValueError: If the input directory does not exist.
+        ValueError: If the input directory does not exist or if sizes are incorrect.
         IOError: If there is an error generating image data from the input directory
-        or saving data to the output file.
-
+                 or saving data to the output file.
     The function performs the following steps:
     1. Verifies that the input directory exists.
     2. Initializes an image data generator with rescaling.
     3. Generates image data from the input directory.
     4. Iterates over the dataset and collects images, labels, and filenames.
-    5. Stores the collected data in a dictionary.
-    6. Saves the dictionary to the specified output file using pickle serialization.
+    5. Splits the data into training, validation, and test sets.
+    6. Stores the collected data in a dictionary.
+    7. Saves the dictionary to the specified output file using pickle serialization.
     """
-
     if not input_dir.exists():
         msg = f"The input directory {input_dir} does not exist."
         raise ValueError(msg)
+    if not 0 < train_size < 1:
+        raise ValueError("Training set percentage must be a value between 0 and 1.")
+    if not 0 < val_size < 1:
+        raise ValueError("Validation set percentage must be a value between 0 and 1.")
+    if train_size + val_size >= 1:
+        raise ValueError(
+            "The sum of training and validation set percentages cannot exceed or be equal to 1.")
 
+    test_size = 1 - train_size - val_size
     images_generator = tf.keras.preprocessing.image.ImageDataGenerator(rescale=1. / 255)
 
     try:
@@ -64,38 +76,46 @@ def save_dataset(input_dir:Path, output_file:Path) -> None:
         msg = f"Error generating image data from {input_dir}: {e}"
[…]

### 2024-09-18 tests/test_dataset_utils.py

@@ -19,21 +19,34 @@ class SaveDatasetTestCase(unittest.TestCase):
         self.tempdir.cleanup()
 
     def test_save_dataset(self):
-        input_dir = Path(self.tempdir.name)
-        output_file = Path(os.path.join(self.tempdir.name, "output.pkl"))
+        input_dir = Path("./images")
+        output_file = Path(os.path.join(self.tempdir.name, "test_save_dataset.pkl"))
         save_dataset(input_dir, output_file)
 
+        assert os.path.exists(output_file)
+
     def test_load_dataset(self):
-        input_file = Path("imagenet_224.pkl")
-        (train_images, train_labels, batch_image_files) = load_dataset(input_file)
+        input_file = Path("test_load_dataset.pkl")
+        (
+            train_images,
+            train_labels,
+            val_images,
+            val_labels,
+            test_images,
+            test_labels,
+            filenames) = load_dataset(input_file)
 
-        assert len(train_images) > 0
-        assert train_images.shape == (32, 256, 256, 3)
+        assert len(train_images) == 28
+        assert len(test_images) == 8
+        assert len(val_images) == 4
 
-        assert len(train_labels) > 0
-        assert train_labels.shape == (32, 10)
+        assert len(train_labels) == 28
+        assert len(test_labels) == 8
+        assert len(val_labels) == 4
 
-        assert len(batch_image_files) == 32
+        assert len(filenames["train"]) == 28
+        assert len(filenames["test"]) == 8
+        assert len(filenames["val"]) == 4
 
     def test_dir_not_existing(self):
         input_dir = Path("nonexistent")

### 2024-09-18 project/utils/dataset_utils.py
Refactor dataset utility functions and improve error messages
Refactor variable names in `save_dataset` for clarity and consistency. Improved error message readability and updated the `load_dataset` docstring to align with the function's behavior. Adjusted test case paths to use a standardized input directory.

@@ -25,10 +25,9 @@ Attributes:
 import argparse
 import pickle
 from pathlib import Path
-from typing import List, Tuple
-from sklearn.model_selection import train_test_split
 
 import tensorflow as tf
+from sklearn.model_selection import train_test_split
 
 
 def save_dataset(
@@ -60,12 +59,15 @@ def save_dataset(
         msg = f"The input directory {input_dir} does not exist."
         raise ValueError(msg)
     if not 0 < train_size < 1:
-        raise ValueError("Training set percentage must be a value between 0 and 1.")
+        msg = "Training set percentage must be a value between 0 and 1."
+        raise ValueError(msg)
     if not 0 < val_size < 1:
-        raise ValueError("Validation set percentage must be a value between 0 and 1.")
+        msg = "Validation set percentage must be a value between 0 and 1."
+        raise ValueError(msg)
     if train_size + val_size >= 1:
-        raise ValueError(
-            "The sum of training and validation set percentages cannot exceed or be equal to 1.")
+        msg = ("The sum of training and validation set percentages cannot exceed or "
+               "be equal to 1.")
+        raise ValueError(msg)
 
     test_size = 1 - train_size - val_size
     images_generator = tf.keras.preprocessing.image.ImageDataGenerator(rescale=1. / 255)
@@ -89,50 +91,65 @@ def save_dataset(
         except StopIteration:
             break
 
-    X_train, X_temp, y_train, y_temp, filenames_train, filenames_temp = train_test_split(
-        all_images, all_labels, all_filenames, train_size=train_size, random_state=42)
+    (x_train,
+     x_temp,
+     y_train,
+     y_temp,
+     filenames_train,
+     filenames_temp) = train_test_split(
+        all_images,
+        all_labels,
+        all_filenames,
+        train_size=train_size,
+        random_state=42)
 
-    X_val, X_test, y_val, y_test, filenames_test, filenames_val = train_test_split(
-        X_temp, y_temp, filenames_temp, test_size=(test_size / (test_size + val_size)),
+    (x_val,
+     x_test,
+     y_val,
+     y_test,
+     filenames_test,
+     filenames_val) = train_test_split(
+        x_temp,
+        y_temp,
+        filenames_temp,
+        test_size=(test_size / (test_size + val_size)),
         random_state=42)
 
     data_to_save = dict(
-        train_images=X_train,
+        train_images=x_train,
         train_labels=y_train,
-        val_images=X_val,
+        val_images=x_val,
         val_labels=y_val,
-        test_images=X_test,
[…]

### 2024-09-18 tests/test_dataset_utils.py

@@ -56,7 +56,7 @@ class SaveDatasetTestCase(unittest.TestCase):
             save_dataset(input_dir, output_file)
 
     def test_wrong_output_file_path(self):
-        input_dir = Path(self.tempdir.name)
+        input_dir = Path("./images")
         output_file = Path(os.path.join(self.tempdir.name, "nonexistent", "output.pkl"))
 
         with pytest.raises(IOError):

### 2024-09-18 project/utils/dataset_utils.py
Enhance image data generator with augmentation settings
Added shear, zoom, and horizontal flip augmentations to the `ImageDataGenerator` configuration. Adjusted the flow_from_directory method to include a batch size and class mode for better handling of image input.

@@ -28,7 +28,7 @@ from pathlib import Path
 
 import tensorflow as tf
 from sklearn.model_selection import train_test_split
-
+from tensorflow.keras.preprocessing.image import ImageDataGenerator
 
 def save_dataset(
         input_dir: Path,
@@ -69,11 +69,18 @@ def save_dataset(
                "be equal to 1.")
         raise ValueError(msg)
 
-    test_size = 1 - train_size - val_size
-    images_generator = tf.keras.preprocessing.image.ImageDataGenerator(rescale=1. / 255)
+    images_generator = ImageDataGenerator(
+        rescale = 1./255,
+        shear_range = 0.2,
+        zoom_range = 0.2,
+        horizontal_flip = True)
 
     try:
-        generator = images_generator.flow_from_directory(input_dir)
+        generator = images_generator.flow_from_directory(
+            input_dir,
+            batch_size=32,
+            class_mode="binary")
+
     except Exception as e:
         msg = f"Error generating image data from {input_dir}: {e}"
         raise IOError(msg) from e
@@ -90,6 +97,8 @@ def save_dataset(
             all_filenames.extend(batch_filenames)
         except StopIteration:
             break
+
+    test_size = 1 - train_size - val_size
 
     (x_train,
      x_temp,

### 2024-09-18 tests/test_dataset_utils.py
Add shape validation for images in dataset tests
Ensure that all images in the train, val, and test sets have the shape (224, 224, 3). This confirms the images meet the expected dimensions, adding an extra layer of validation to the dataset utility tests.

@@ -48,6 +48,9 @@ class SaveDatasetTestCase(unittest.TestCase):
         assert len(filenames["test"]) == 8
         assert len(filenames["val"]) == 4
 
+        for image in train_images + val_images + test_images:
+            assert image.shape == (224, 224, 3)
+
     def test_dir_not_existing(self):
         input_dir = Path("nonexistent")
         output_file = Path(os.path.join(self.tempdir.name, "output.pkl"))

### 2024-09-18 project/utils/dataset_utils.py
Refactor `save_dataset` function to include image resizing
Added an optional `image_target_size` parameter to the `save_dataset` function to specify the target size for image resizing. Updated the docstring to reflect this change and provided clear parameter descriptions and exception handling details.

@@ -25,35 +25,37 @@ Attributes:
 import argparse
 import pickle
 from pathlib import Path
+from typing import Tuple
 
-import tensorflow as tf
 from sklearn.model_selection import train_test_split
 from tensorflow.keras.preprocessing.image import ImageDataGenerator
+
 
 def save_dataset(
         input_dir: Path,
         output_file: Path,
+        image_target_size: Tuple[int, int]=(224, 224),
         train_size=0.7,
         val_size=0.2) -> None:
     """
-    Save a dataset of images and their corresponding labels to a file.
-    Args:
-        input_dir (Path): The directory containing the image dataset.
-        output_file (Path): The file path where the dataset will be saved.
-        train_size (float): Proportion of the dataset to include in the training set.
-        val_size (float): Proportion of the dataset to include in the validation set.
-    Raises:
-        ValueError: If the input directory does not exist or if sizes are incorrect.
-        IOError: If there is an error generating image data from the input directory
-                 or saving data to the output file.
-    The function performs the following steps:
-    1. Verifies that the input directory exists.
-    2. Initializes an image data generator with rescaling.
-    3. Generates image data from the input directory.
-    4. Iterates over the dataset and collects images, labels, and filenames.
-    5. Splits the data into training, validation, and test sets.
-    6. Stores the collected data in a dictionary.
-    7. Saves the dictionary to the specified output file using pickle serialization.
+        Save images from a directory to a file in a structured format, while splitting
+        them into training, validation, and test sets.
+
+        Parameters:
+        input_dir (Path): The directory containing the image files.
+        output_file (Path): The file to save the structured dataset.
+        image_target_size (Tuple[int, int], optional): The target size for resizing
+        images. Defaults to (224, 224).
+        train_size (float, optional): The proportion of images to be used for training.
+        Defaults to 0.7.
+        val_size (float, optional): The proportion of images to be used for validation.
+        Defaults to 0.2.
+
+        Raises:
+        ValueError: If the input directory does not exist, or if train_size and val_size
+        are not within the range (0, 1), or if their sum is >= 1.
+        IOError: If there is an error generating image data from the input directory,
+        or saving the data to the output file.
     """
     if not input_dir.exists():
         msg = f"The input directory {input_dir} does not exist."
@@ -79,7 +81,8 @@ def save_dataset(
         generator = images_generator.flow_from_directory(
             input_dir,
             batch_size=32,
-            class_mode="binary")
[…]

### 2024-09-18 project/utils/image_utils.py
Add utility for creating train, test, validation sets
Create a new function `create_train_test_validation_sets` in `image_utils.py` to organize files into training, validation, and testing directories. Added unit tests in `test_image_utils.py` to verify the creation of these directories.

@@ -1,0 +1,23 @@
+import os.path
+from pathlib import Path
+
+def create_train_test_validation_sets(
+        input_dir:Path,
+        output_dir:Path,
+        filter) -> None:
+
+    files = input_dir.rglob(filter)
+
+    training_set_folder_path = os.path.join(output_dir,"train")
+    validation_set_folder_path = os.path.join(output_dir, "val")
+    test_set_folder_path = os.path.join(output_dir,"test")
+
+    if not os.path.exists(training_set_folder_path):
+        os.makedirs(training_set_folder_path)
+
+    if not os.path.exists(test_set_folder_path):
+        os.makedirs(test_set_folder_path)
+
+    if not os.path.exists(validation_set_folder_path):
+        os.makedirs(validation_set_folder_path)
+

### 2024-09-18 tests/test_image_utils.py

@@ -1,0 +1,27 @@
+import unittest
+import os
+
+from project.utils.image_utils import create_train_test_validation_sets
+from pathlib import Path
+
+class TestImageUtils(unittest.TestCase):
+
+    def setUp(self):
+        pass
+
+    def tearDown(self):
+        pass
+
+    def test_create_train_test_validation_sets(self):
+        input_dir = Path("./images224")
+        output_dir = Path("test_create_train_test_validation_sets")
+        filter = "*.jpeg"
+
+        create_train_test_validation_sets(input_dir, output_dir, filter)
+
+        assert os.path.exists(os.path.join(output_dir, "train"))
+        assert os.path.exists(os.path.join(output_dir, "val"))
+        assert os.path.exists(os.path.join(output_dir, "test"))
+
+if __name__ == "__main__":
+    unittest.main()

### 2024-09-18 project/utils/image_utils.py
Simplify directory creation logic
Refactor the code to remove redundancy by using a loop to create the 'train', 'val', and 'test' directories. This new approach makes the code cleaner and more maintainable.

@@ -7,17 +7,10 @@ def create_train_test_validation_sets(
         filter) -> None:
 
     files = input_dir.rglob(filter)
+    output_folders = ["train", "val", "test"]
 
-    training_set_folder_path = os.path.join(output_dir,"train")
-    validation_set_folder_path = os.path.join(output_dir, "val")
-    test_set_folder_path = os.path.join(output_dir,"test")
-
-    if not os.path.exists(training_set_folder_path):
-        os.makedirs(training_set_folder_path)
-
-    if not os.path.exists(test_set_folder_path):
-        os.makedirs(test_set_folder_path)
-
-    if not os.path.exists(validation_set_folder_path):
-        os.makedirs(validation_set_folder_path)
+    for output_folder in output_folders:
+        full_path = os.path.join(output_dir, output_folder)
+        if not os.path.exists(full_path):
+            os.makedirs(full_path)
 

### 2024-09-18 project/utils/image_utils.py
Add train percentage parameter to dataset creation function
Updated `create_train_test_validation_sets` to accept a `train_pct` parameter. Added folder path setup and existence checks. Modified test cases to align with updated function signature and new folder creation logic.

@@ -4,13 +4,17 @@ from pathlib import Path
 def create_train_test_validation_sets(
         input_dir:Path,
         output_dir:Path,
-        filter) -> None:
+        filter,
+        train_pct:float) -> None:
+
+    set_folder_paths = dict(
+        train=os.path.join(output_dir, "train"),
+        val=os.path.join(output_dir, "val"),
+        test=os.path.join(output_dir, "test")
+    )
+
+    for folder_path in set_folder_paths.values():
+        if not os.path.exists(folder_path):
+            os.makedirs(folder_path)
 
     files = input_dir.rglob(filter)
-    output_folders = ["train", "val", "test"]
-
-    for output_folder in output_folders:
-        full_path = os.path.join(output_dir, output_folder)
-        if not os.path.exists(full_path):
-            os.makedirs(full_path)
-

### 2024-09-18 tests/test_image_utils.py

@@ -17,11 +17,17 @@ class TestImageUtils(unittest.TestCase):
         output_dir = Path("test_create_train_test_validation_sets")
         filter = "*.jpeg"
 
-        create_train_test_validation_sets(input_dir, output_dir, filter)
+        create_train_test_validation_sets(
+            input_dir,
+            output_dir,
+            filter,
+            0.7)
 
-        assert os.path.exists(os.path.join(output_dir, "train"))
-        assert os.path.exists(os.path.join(output_dir, "val"))
-        assert os.path.exists(os.path.join(output_dir, "test"))
+        train_folder = Path(os.path.join(output_dir, "train"))
+        files = list(train_folder.rglob(filter))
+        # assert len(files) == 14
+
+
 
 if __name__ == "__main__":
     unittest.main()

### 2024-09-18 project/utils/image_utils.py
Refactor folder creation logic in image_utils.py
Modularize folder creation by introducing two helper functions: `_create_output_folder` and `_create_image_class_folder`. Update unit tests to verify the existence of the necessary folders for training, validation, and testing within the specified directories.

@@ -1,20 +1,36 @@
 import os.path
 from pathlib import Path
 
+from tensorboard.summary.v1 import image
+
+
 def create_train_test_validation_sets(
         input_dir:Path,
         output_dir:Path,
         filter,
         train_pct:float) -> None:
 
-    set_folder_paths = dict(
-        train=os.path.join(output_dir, "train"),
-        val=os.path.join(output_dir, "val"),
-        test=os.path.join(output_dir, "test")
-    )
+    output_folders = [_create_output_folder(output_dir, f) for f
+                      in ["train", "test", "val"]]
 
-    for folder_path in set_folder_paths.values():
-        if not os.path.exists(folder_path):
-            os.makedirs(folder_path)
+    image_class_folders = [_create_image_class_folder(of, image_class)
+                           for of in output_folders
+                           for image_class in os.listdir(input_dir)]
 
-    files = input_dir.rglob(filter)
+
+
+def _create_output_folder(root_path:Path, folder_name:str) -> Path:
+    output_path = os.path.join(root_path, folder_name)
+
+    if not os.path.exists(output_path):
+        os.makedirs(output_path)
+
+    return output_path
+
+def _create_image_class_folder(root_path:Path, image_class:str) -> Path:
+    image_class_path = os.path.join(root_path, image_class)
+
+    if not os.path.exists(image_class_path):
+        os.makedirs(image_class_path)
+
+    return image_class_path

### 2024-09-18 tests/test_image_utils.py

@@ -15,6 +15,10 @@ class TestImageUtils(unittest.TestCase):
     def test_create_train_test_validation_sets(self):
         input_dir = Path("./images224")
         output_dir = Path("test_create_train_test_validation_sets")
+
+        if not os.path.exists(output_dir):
+            os.makedirs(output_dir)
+
         filter = "*.jpeg"
 
         create_train_test_validation_sets(
@@ -23,9 +27,16 @@ class TestImageUtils(unittest.TestCase):
             filter,
             0.7)
 
-        train_folder = Path(os.path.join(output_dir, "train"))
-        files = list(train_folder.rglob(filter))
-        # assert len(files) == 14
+        train_folder = Path(os.path.join(output_dir, "train", "bird"))
+        val_folder = Path(os.path.join(output_dir, "val", "bird"))
+        test_folder = Path(os.path.join(output_dir, "test", "bird"))
+
+        assert os.path.exists(train_folder)
+        assert os.path.exists(val_folder)
+        assert os.path.exists(test_folder)
+
+        # train_file_count = list(train_folder.rglob(filter))
+        # assert len(train_file_count) == 14
 
 
 

### 2024-09-19 project/utils/image_utils.py
Add random shuffling and copying of training files
This update includes the addition of random shuffling for images and copying them into their respective training folders. The `_create_image_class_folder` function has been renamed to `_create_image_class_output_folder` for clarity, and test cases have been updated to validate the corrected file count.

@@ -1,4 +1,7 @@
 import os.path
+import random
+import shutil
+
 from pathlib import Path
 
 from tensorboard.summary.v1 import image
@@ -13,10 +16,26 @@ def create_train_test_validation_sets(
     output_folders = [_create_output_folder(output_dir, f) for f
                       in ["train", "test", "val"]]
 
-    image_class_folders = [_create_image_class_folder(of, image_class)
-                           for of in output_folders
-                           for image_class in os.listdir(input_dir)]
+    source_image_class_folders = os.listdir(input_dir)
 
+    image_class_output_folders = [_create_image_class_output_folder(of, image_class)
+                                  for of in output_folders
+                                  for image_class in source_image_class_folders]
+
+    for source_image_class_folder in source_image_class_folders:
+        source_image_class_folder_full_path = os.path.join(input_dir, source_image_class_folder)
+
+        files =[os.path.join(source_image_class_folder_full_path, fn)
+                for fn in os.listdir(source_image_class_folder_full_path)]
+
+        [random.shuffle(files) for _ in range(0, 3)]
+
+        file_count = len(files)
+        train_file_count = int(file_count * train_pct)
+        train_files = files[:train_file_count]
+
+        for file in train_files:
+            shutil.copy2(file, image_class_output_folders[0])
 
 
 def _create_output_folder(root_path:Path, folder_name:str) -> Path:
@@ -27,7 +46,7 @@ def _create_output_folder(root_path:Path, folder_name:str) -> Path:
 
     return output_path
 
-def _create_image_class_folder(root_path:Path, image_class:str) -> Path:
+def _create_image_class_output_folder(root_path:Path, image_class:str) -> Path:
     image_class_path = os.path.join(root_path, image_class)
 
     if not os.path.exists(image_class_path):

### 2024-09-19 tests/test_image_utils.py

@@ -35,8 +35,8 @@ class TestImageUtils(unittest.TestCase):
         assert os.path.exists(val_folder)
         assert os.path.exists(test_folder)
 
-        # train_file_count = list(train_folder.rglob(filter))
-        # assert len(train_file_count) == 14
+        train_file_count = list(train_folder.rglob(filter))
+        assert len(train_file_count) == 14
 
 
 

### 2024-09-19 project/utils/image_utils.py
Add support for testing file splitting in image manager.
Enhanced the image manager to include a test percent parameter for splitting files into a test set. Updated corresponding unit tests to verify the correct allocation of files into train, test, and validation folders.

@@ -11,7 +11,8 @@ def create_train_test_validation_sets(
         input_dir:Path,
         output_dir:Path,
         filter,
-        train_pct:float) -> None:
+        train_pct:float,
+        test_pct:float) -> None:
 
     output_folders = [_create_output_folder(output_dir, f) for f
                       in ["train", "test", "val"]]
@@ -36,6 +37,12 @@ def create_train_test_validation_sets(
 
         for file in train_files:
             shutil.copy2(file, image_class_output_folders[0])
+
+        test_file_count = int(file_count * test_pct)
+        test_files = files[train_file_count:train_file_count + test_file_count]
+
+        for file in test_files:
+            shutil.copy2(file, image_class_output_folders[1])
 
 
 def _create_output_folder(root_path:Path, folder_name:str) -> Path:

### 2024-09-19 tests/test_image_utils.py

@@ -25,11 +25,12 @@ class TestImageUtils(unittest.TestCase):
             input_dir,
             output_dir,
             filter,
-            0.7)
+            0.7,
+        0.2)
 
         train_folder = Path(os.path.join(output_dir, "train", "bird"))
-        val_folder = Path(os.path.join(output_dir, "val", "bird"))
         test_folder = Path(os.path.join(output_dir, "test", "bird"))
+        val_folder = Path(os.path.join(output_dir, "val", "bird"))
 
         assert os.path.exists(train_folder)
         assert os.path.exists(val_folder)
@@ -37,6 +38,9 @@ class TestImageUtils(unittest.TestCase):
 
         train_file_count = list(train_folder.rglob(filter))
         assert len(train_file_count) == 14
+
+        test_file_count = list(test_folder.rglob(filter))
+        assert len(test_file_count) == 4
 
 
 

### 2024-09-19 tests/test_image_utils.py
Refactor test setup and teardown in test_image_utils.py
Updated `setUp` and `tearDown` methods to initialize and clean test directories. This ensures test files are handled consistently and cleanup is properly executed after tests run.

@@ -1,5 +1,6 @@
 import unittest
 import os
+import shutil
 
 from project.utils.image_utils import create_train_test_validation_sets
 from pathlib import Path
@@ -7,30 +8,28 @@ from pathlib import Path
 class TestImageUtils(unittest.TestCase):
 
     def setUp(self):
-        pass
+        self._input_dir = Path("./images224")
+        self._output_dir = Path("test_create_train_test_validation_sets")
 
     def tearDown(self):
-        pass
+        shutil.rmtree(self._output_dir)
 
     def test_create_train_test_validation_sets(self):
-        input_dir = Path("./images224")
-        output_dir = Path("test_create_train_test_validation_sets")
-
-        if not os.path.exists(output_dir):
-            os.makedirs(output_dir)
+        if not os.path.exists(self._output_dir):
+            os.makedirs(self._output_dir)
 
         filter = "*.jpeg"
 
         create_train_test_validation_sets(
-            input_dir,
-            output_dir,
+            self._input_dir,
+            self._output_dir,
             filter,
             0.7,
         0.2)
 
-        train_folder = Path(os.path.join(output_dir, "train", "bird"))
-        test_folder = Path(os.path.join(output_dir, "test", "bird"))
-        val_folder = Path(os.path.join(output_dir, "val", "bird"))
+        train_folder = Path(os.path.join(self._output_dir, "train", "bird"))
+        test_folder = Path(os.path.join(self._output_dir, "test", "bird"))
+        val_folder = Path(os.path.join(self._output_dir, "val", "bird"))
 
         assert os.path.exists(train_folder)
         assert os.path.exists(val_folder)

### 2024-09-19 project/utils/image_utils.py
Add validation split in image_utils and update tests
Enhanced `image_utils.py` to include validation file splitting and copying, ensuring files are distributed across train, test, and validation folders. Updated `test_image_utils.py` to verify correct validation file count, maintaining test coverage and functionality.

@@ -44,6 +44,11 @@ def create_train_test_validation_sets(
         for file in test_files:
             shutil.copy2(file, image_class_output_folders[1])
 
+        validation_file_count = file_count - train_file_count - test_file_count
+        validation_files = files[train_file_count + test_file_count:]
+
+        for file in validation_files:
+            shutil.copy2(file, image_class_output_folders[2])
 
 def _create_output_folder(root_path:Path, folder_name:str) -> Path:
     output_path = os.path.join(root_path, folder_name)

### 2024-09-19 tests/test_image_utils.py

@@ -41,7 +41,8 @@ class TestImageUtils(unittest.TestCase):
         test_file_count = list(test_folder.rglob(filter))
         assert len(test_file_count) == 4
 
-
+        val_file_count = list(val_folder.rglob(filter))
+        assert len(val_file_count) == 2
 
 if __name__ == "__main__":
     unittest.main()

### 2024-09-19 project/utils/image_utils.py
Remove unused imports and reformat code for clarity
Removed unused imports and slight reformatting of lines to enhance readability in `image_utils.py`. Also reorganized imports in the test file to comply with PEP 8 standards.

@@ -1,10 +1,7 @@
 import os.path
 import random
 import shutil
-
 from pathlib import Path
-
-from tensorboard.summary.v1 import image
 
 
 def create_train_test_validation_sets(
@@ -24,7 +21,10 @@ def create_train_test_validation_sets(
                                   for image_class in source_image_class_folders]
 
     for source_image_class_folder in source_image_class_folders:
-        source_image_class_folder_full_path = os.path.join(input_dir, source_image_class_folder)
+
+        source_image_class_folder_full_path = os.path.join(
+            input_dir,
+            source_image_class_folder)
 
         files =[os.path.join(source_image_class_folder_full_path, fn)
                 for fn in os.listdir(source_image_class_folder_full_path)]
@@ -44,7 +44,6 @@ def create_train_test_validation_sets(
         for file in test_files:
             shutil.copy2(file, image_class_output_folders[1])
 
-        validation_file_count = file_count - train_file_count - test_file_count
         validation_files = files[train_file_count + test_file_count:]
 
         for file in validation_files:

### 2024-09-19 tests/test_image_utils.py

@@ -1,9 +1,10 @@
-import unittest
 import os
 import shutil
+import unittest
+from pathlib import Path
 
 from project.utils.image_utils import create_train_test_validation_sets
-from pathlib import Path
+
 
 class TestImageUtils(unittest.TestCase):
 

### 2024-09-19 project/utils/image_utils.py
Refactor dataset partitioning logic and update tests
Refactor `create_train_test_validation_sets` to improve readability and maintainability by introducing helper functions `_get_files_from_folder` and `_copy_files`. Update constants and variable names for clarity. Modify tests accordingly to match the updated function signature and add constants for split ratios and file counts.

@@ -3,51 +3,54 @@ import random
 import shutil
 from pathlib import Path
 
+TRAIN_FOLDER = "train"
+TEST_FOLDER = "test"
+VAL_FOLDER = "val"
+SHUFFLE_ROUNDS = 3
+
 
 def create_train_test_validation_sets(
-        input_dir:Path,
-        output_dir:Path,
-        filter,
-        train_pct:float,
-        test_pct:float) -> None:
+        input_dir: Path,
+        output_dir: Path,
+        file_filter,
+        train_percentage: float,
+        test_percentage: float) -> None:
+    output_folders = [_create_output_folder(output_dir, folder_name)
+                      for folder_name in [TRAIN_FOLDER, TEST_FOLDER, VAL_FOLDER]]
 
-    output_folders = [_create_output_folder(output_dir, f) for f
-                      in ["train", "test", "val"]]
+    image_class_folders = os.listdir(input_dir)
 
-    source_image_class_folders = os.listdir(input_dir)
+    image_class_output_folders = [
+        _create_image_class_output_folder(output_folder, image_class)
+        for output_folder in output_folders
+        for image_class in image_class_folders
+    ]
 
-    image_class_output_folders = [_create_image_class_output_folder(of, image_class)
-                                  for of in output_folders
-                                  for image_class in source_image_class_folders]
+    for source_image_class_folder in image_class_folders:
+        source_folder_path = os.path.join(input_dir, source_image_class_folder)
+        files = _get_files_from_folder(source_folder_path)
+        random.shuffle(files)
 
-    for source_image_class_folder in source_image_class_folders:
+        train_count = int(len(files) * train_percentage)
+        test_count = int(len(files) * test_percentage)
 
-        source_image_class_folder_full_path = os.path.join(
-            input_dir,
-            source_image_class_folder)
+        train_files = files[:train_count]
+        _copy_files(train_files, image_class_output_folders[0])
 
-        files =[os.path.join(source_image_class_folder_full_path, fn)
-                for fn in os.listdir(source_image_class_folder_full_path)]
+        test_files = files[train_count:train_count + test_count]
+        _copy_files(test_files, image_class_output_folders[1])
 
-        [random.shuffle(files) for _ in range(0, 3)]
+        validation_files = files[train_count + test_count:]
+        _copy_files(validation_files, image_class_output_folders[2])
 
-        file_count = len(files)
-        train_file_count = int(file_count * train_pct)
[…]

### 2024-09-19 tests/test_image_utils.py

@@ -5,9 +5,15 @@ from pathlib import Path
 
 from project.utils.image_utils import create_train_test_validation_sets
 
+TRAIN_SPLIT = 0.7
+TEST_SPLIT = 0.2
+TRAIN_FILE_COUNT = 14
+TEST_FILE_COUNT = 4
+VAL_FILE_COUNT = 2
+JPEG_FILTER = "*.jpeg"
+
 
 class TestImageUtils(unittest.TestCase):
-
     def setUp(self):
         self._input_dir = Path("./images224")
         self._output_dir = Path("test_create_train_test_validation_sets")
@@ -15,18 +21,20 @@ class TestImageUtils(unittest.TestCase):
     def tearDown(self):
         shutil.rmtree(self._output_dir)
 
-    def test_create_train_test_validation_sets(self):
+    def create_output_dir_if_not_exists(self):
         if not os.path.exists(self._output_dir):
             os.makedirs(self._output_dir)
 
-        filter = "*.jpeg"
+    def test_create_train_test_validation_sets(self):
+        self.create_output_dir_if_not_exists()
 
         create_train_test_validation_sets(
-            self._input_dir,
-            self._output_dir,
-            filter,
-            0.7,
-        0.2)
+            input_dir=self._input_dir,
+            output_dir=self._output_dir,
+            file_filter=JPEG_FILTER,
+            train_percentage=TRAIN_SPLIT,
+            test_percentage=TEST_SPLIT,
+        )
 
         train_folder = Path(os.path.join(self._output_dir, "train", "bird"))
         test_folder = Path(os.path.join(self._output_dir, "test", "bird"))
@@ -36,14 +44,14 @@ class TestImageUtils(unittest.TestCase):
         assert os.path.exists(val_folder)
         assert os.path.exists(test_folder)
 
-        train_file_count = list(train_folder.rglob(filter))
-        assert len(train_file_count) == 14
+        train_file_count = list(train_folder.rglob(JPEG_FILTER))
+        assert len(train_file_count) == TRAIN_FILE_COUNT
 
-        test_file_count = list(test_folder.rglob(filter))
-        assert len(test_file_count) == 4
+        test_file_count = list(test_folder.rglob(JPEG_FILTER))
+        assert len(test_file_count) == TEST_FILE_COUNT
 
-        val_file_count = list(val_folder.rglob(filter))
-        assert len(val_file_count) == 2
[…]

### 2024-09-19 project/utils/image_utils.py
Set default values for train and test split percentages
Default values for `train_percentage` and `test_percentage` in `create_train_test_validation_sets` function have been added to `image_utils.py`. Corresponding adjustments were made in `test_image_utils.py` to reflect this change.

@@ -13,8 +13,8 @@ def create_train_test_validation_sets(
         input_dir: Path,
         output_dir: Path,
         file_filter,
-        train_percentage: float,
-        test_percentage: float) -> None:
+        train_percentage: float=0.7,
+        test_percentage: float=0.2) -> None:
     output_folders = [_create_output_folder(output_dir, folder_name)
                       for folder_name in [TRAIN_FOLDER, TEST_FOLDER, VAL_FOLDER]]
 

### 2024-09-19 tests/test_image_utils.py

@@ -5,8 +5,6 @@ from pathlib import Path
 
 from project.utils.image_utils import create_train_test_validation_sets
 
-TRAIN_SPLIT = 0.7
-TEST_SPLIT = 0.2
 TRAIN_FILE_COUNT = 14
 TEST_FILE_COUNT = 4
 VAL_FILE_COUNT = 2
@@ -32,8 +30,6 @@ class TestImageUtils(unittest.TestCase):
             input_dir=self._input_dir,
             output_dir=self._output_dir,
             file_filter=JPEG_FILTER,
-            train_percentage=TRAIN_SPLIT,
-            test_percentage=TEST_SPLIT,
         )
 
         train_folder = Path(os.path.join(self._output_dir, "train", "bird"))

### 2024-09-19 project/utils/image_utils.py
Set default file filter in create_train_test_validation_sets
Add a default value of "*.jpeg" for the file_filter parameter to streamline function calls. Updated the corresponding test case to use the default filter, reducing redundancy in test configuration.

@@ -12,7 +12,7 @@ SHUFFLE_ROUNDS = 3
 def create_train_test_validation_sets(
         input_dir: Path,
         output_dir: Path,
-        file_filter,
+        file_filter="*.jpeg",
         train_percentage: float=0.7,
         test_percentage: float=0.2) -> None:
     output_folders = [_create_output_folder(output_dir, folder_name)

### 2024-09-19 tests/test_image_utils.py

@@ -26,11 +26,7 @@ class TestImageUtils(unittest.TestCase):
     def test_create_train_test_validation_sets(self):
         self.create_output_dir_if_not_exists()
 
-        create_train_test_validation_sets(
-            input_dir=self._input_dir,
-            output_dir=self._output_dir,
-            file_filter=JPEG_FILTER,
-        )
+        create_train_test_validation_sets(self._input_dir, self._output_dir)
 
         train_folder = Path(os.path.join(self._output_dir, "train", "bird"))
         test_folder = Path(os.path.join(self._output_dir, "test", "bird"))

### 2024-09-19 project/utils/image_utils.py
Refactor folder creation in dataset splitting function
Simplify the process of generating output folders dynamically within the loop that processes each image class. This change reduces unnecessary complexity and ensures proper folder creation for training, testing, and validation sets.

@@ -8,7 +8,6 @@ TEST_FOLDER = "test"
 VAL_FOLDER = "val"
 SHUFFLE_ROUNDS = 3
 
-
 def create_train_test_validation_sets(
         input_dir: Path,
         output_dir: Path,
@@ -20,14 +19,8 @@ def create_train_test_validation_sets(
 
     image_class_folders = os.listdir(input_dir)
 
-    image_class_output_folders = [
-        _create_image_class_output_folder(output_folder, image_class)
-        for output_folder in output_folders
-        for image_class in image_class_folders
-    ]
-
-    for source_image_class_folder in image_class_folders:
-        source_folder_path = os.path.join(input_dir, source_image_class_folder)
+    for image_class_folder in image_class_folders:
+        source_folder_path = os.path.join(input_dir, image_class_folder)
         files = _get_files_from_folder(source_folder_path)
         random.shuffle(files)
 
@@ -35,14 +28,26 @@ def create_train_test_validation_sets(
         test_count = int(len(files) * test_percentage)
 
         train_files = files[:train_count]
-        _copy_files(train_files, image_class_output_folders[0])
-
         test_files = files[train_count:train_count + test_count]
-        _copy_files(test_files, image_class_output_folders[1])
-
         validation_files = files[train_count + test_count:]
-        _copy_files(validation_files, image_class_output_folders[2])
 
+        train_folder = _create_image_class_output_folder(
+            output_folders[0],
+            image_class_folder)
+
+        _copy_files(train_files, train_folder)
+
+        test_folder = _create_image_class_output_folder(
+            output_folders[1],
+            image_class_folder)
+
+        _copy_files(test_files, test_folder)
+
+        validation_folder = _create_image_class_output_folder(
+            output_folders[2],
+            image_class_folder)
+
+        _copy_files(validation_files, validation_folder)
 
 def _get_files_from_folder(folder_path: str) -> list:
     return [os.path.join(folder_path, filename) for filename in os.listdir(folder_path)]

### 2024-09-19 tests/test_image_utils.py
Refactor test with parameterized.expand
Refactor `test_create_train_test_validation_sets` to use `parameterized.expand` for testing multiple image classes. Updated assertions to dynamically check expected file counts for training, testing, and validation sets. Added helper method `_get_file_count` to streamline folder file count logic.

@@ -3,6 +3,8 @@ import shutil
 import unittest
 from pathlib import Path
 
+from parameterized import parameterized
+
 from project.utils.image_utils import create_train_test_validation_sets
 
 TRAIN_FILE_COUNT = 14
@@ -23,27 +25,30 @@ class TestImageUtils(unittest.TestCase):
         if not os.path.exists(self._output_dir):
             os.makedirs(self._output_dir)
 
-    def test_create_train_test_validation_sets(self):
+    @parameterized.expand([
+        ("bird", 14, 4, 2),
+        ("cat", 14, 4, 2),
+    ])
+    def test_create_train_test_validation_sets(
+            self,
+            image_class,
+            expected_train_file_count,
+            expected_test_file_count,
+            expected_validation_file_count):
         self.create_output_dir_if_not_exists()
 
         create_train_test_validation_sets(self._input_dir, self._output_dir)
 
-        train_folder = Path(os.path.join(self._output_dir, "train", "bird"))
-        test_folder = Path(os.path.join(self._output_dir, "test", "bird"))
-        val_folder = Path(os.path.join(self._output_dir, "val", "bird"))
+        train_folder = Path(os.path.join(self._output_dir, "train", image_class))
+        test_folder = Path(os.path.join(self._output_dir, "test", image_class))
+        val_folder = Path(os.path.join(self._output_dir, "val", image_class))
 
-        assert os.path.exists(train_folder)
-        assert os.path.exists(val_folder)
-        assert os.path.exists(test_folder)
+        assert self._get_file_count(train_folder) == expected_train_file_count
+        assert self._get_file_count(test_folder) == expected_test_file_count
+        assert self._get_file_count(val_folder) == expected_validation_file_count
 
-        train_file_count = list(train_folder.rglob(JPEG_FILTER))
-        assert len(train_file_count) == TRAIN_FILE_COUNT
-
-        test_file_count = list(test_folder.rglob(JPEG_FILTER))
-        assert len(test_file_count) == TEST_FILE_COUNT
-
-        val_file_count = list(val_folder.rglob(JPEG_FILTER))
-        assert len(val_file_count) == VAL_FILE_COUNT
+    def _get_file_count(self, folder:Path) -> int:
+        return len(list(folder.rglob(JPEG_FILTER)))
 
 if __name__ == "__main__":
     unittest.main()

### 2024-09-19 project/utils/image_utils.py
Refactor file retrieval logic in image_utils.py
Reorganized the code by extracting a new function `_get_source_files` for clarity. This improves code readability and adheres to single-responsibility principles, making the file retrieval process more maintainable. Also included type annotations to enhance code quality.

@@ -2,6 +2,7 @@ import os.path
 import random
 import shutil
 from pathlib import Path
+from typing import List
 
 TRAIN_FOLDER = "train"
 TEST_FOLDER = "test"
@@ -20,8 +21,7 @@ def create_train_test_validation_sets(
     image_class_folders = os.listdir(input_dir)
 
     for image_class_folder in image_class_folders:
-        source_folder_path = os.path.join(input_dir, image_class_folder)
-        files = _get_files_from_folder(source_folder_path)
+        files = _get_source_files(image_class_folder, input_dir)
         random.shuffle(files)
 
         train_count = int(len(files) * train_percentage)
@@ -49,7 +49,13 @@ def create_train_test_validation_sets(
 
         _copy_files(validation_files, validation_folder)
 
-def _get_files_from_folder(folder_path: str) -> list:
+
+def _get_source_files(image_class_folder:str, input_dir:Path) -> List[str]:
+    source_folder_path = os.path.join(input_dir, image_class_folder)
+    return _get_files_from_folder(source_folder_path)
+
+
+def _get_files_from_folder(folder_path: str) -> List[str]:
     return [os.path.join(folder_path, filename) for filename in os.listdir(folder_path)]
 
 

### 2024-09-19 project/utils/image_utils.py
Refactor file splitting logic in image_utils.py
Move the file splitting logic into a new helper function `_split_files`. This change simplifies the main function and enhances code reusability and readability.

@@ -24,12 +24,10 @@ def create_train_test_validation_sets(
         files = _get_source_files(image_class_folder, input_dir)
         random.shuffle(files)
 
-        train_count = int(len(files) * train_percentage)
-        test_count = int(len(files) * test_percentage)
-
-        train_files = files[:train_count]
-        test_files = files[train_count:train_count + test_count]
-        validation_files = files[train_count + test_count:]
+        test_files, train_files, validation_files = _split_files(
+            files,
+            test_percentage,
+            train_percentage)
 
         train_folder = _create_image_class_output_folder(
             output_folders[0],
@@ -48,6 +46,20 @@ def create_train_test_validation_sets(
             image_class_folder)
 
         _copy_files(validation_files, validation_folder)
+
+
+def _split_files(
+        files:List[str],
+        test_percentage:float,
+        train_percentage:float) -> Tuple[List[str], List[str], List[str]]:
+    train_count = int(len(files) * train_percentage)
+    test_count = int(len(files) * test_percentage)
+
+    train_files = files[:train_count]
+    test_files = files[train_count:train_count + test_count]
+    validation_files = files[train_count + test_count:]
+
+    return test_files, train_files, validation_files
 
 
 def _get_source_files(image_class_folder:str, input_dir:Path) -> List[str]:

### 2024-09-19 project/utils/image_utils.py
Refactor file splitting and copying logic
Reorganize the logic to first split files into training, testing, and validation sets before copying them to their respective output folders. Introduce a helper function `_copy_files_to_output` to reduce redundancy and improve code clarity.

@@ -2,7 +2,7 @@ import os.path
 import random
 import shutil
 from pathlib import Path
-from typing import List
+from typing import List, Tuple
 
 TRAIN_FOLDER = "train"
 TEST_FOLDER = "test"
@@ -24,29 +24,24 @@ def create_train_test_validation_sets(
         files = _get_source_files(image_class_folder, input_dir)
         random.shuffle(files)
 
-        test_files, train_files, validation_files = _split_files(
+        train_files, test_files, validation_files = _split_files(
             files,
             test_percentage,
             train_percentage)
 
-        train_folder = _create_image_class_output_folder(
-            output_folders[0],
-            image_class_folder)
+        _copy_files_to_output(output_folders[0], image_class_folder, train_files)
+        _copy_files_to_output(output_folders[1], image_class_folder, test_files)
+        _copy_files_to_output(output_folders[2], image_class_folder, validation_files)
 
-        _copy_files(train_files, train_folder)
+def _copy_files_to_output(
+        output_folder:Path,
+        image_class_folder:str,
+        files:List[str]) -> None:
+    folder = _create_image_class_output_folder(
+        output_folder,
+        image_class_folder)
 
-        test_folder = _create_image_class_output_folder(
-            output_folders[1],
-            image_class_folder)
-
-        _copy_files(test_files, test_folder)
-
-        validation_folder = _create_image_class_output_folder(
-            output_folders[2],
-            image_class_folder)
-
-        _copy_files(validation_files, validation_folder)
-
+    _copy_files(files, folder)
 
 def _split_files(
         files:List[str],
@@ -59,7 +54,7 @@ def _split_files(
     test_files = files[train_count:train_count + test_count]
     validation_files = files[train_count + test_count:]
 
-    return test_files, train_files, validation_files
+    return train_files, test_files, validation_files
 
 
 def _get_source_files(image_class_folder:str, input_dir:Path) -> List[str]:

### 2024-09-19 project/utils/image_utils.py
Add docstrings and type annotations to image_utils module
Added comprehensive docstrings to all functions in the `image_utils` module to improve code documentation and readability. Also formatted type annotations for better clarity and consistency.

@@ -1,3 +1,24 @@
+"""
+Module: image_utils
+
+This module contains utilities for creating train, test, and validation sets
+from a directory of images. The sets are created based on given percentages
+for training and testing, with the remainder being used for validation.
+
+Functions:
+    create_train_test_validation_sets: Main function to create the dataset splits.
+    _copy_files_to_output: Helper function to copy files to the output directory.
+    _split_files: Helper function to split the files into train, test, and
+    validation sets.
+    _get_source_files: Helper function to retrieve source files from a directory.
+    _get_files_from_folder: Helper function to get files from a specific folder.
+    _copy_files: Helper function to copy files to the specified output folder.
+    _create_output_folder: Helper function to create an output directory if it
+    doesn't exist.
+    _create_image_class_output_folder: Helper function to create sub-directories within
+    the output directory.
+"""
+
 import os.path
 import random
 import shutil
@@ -9,12 +30,31 @@ TEST_FOLDER = "test"
 VAL_FOLDER = "val"
 SHUFFLE_ROUNDS = 3
 
+
 def create_train_test_validation_sets(
         input_dir: Path,
         output_dir: Path,
         file_filter="*.jpeg",
-        train_percentage: float=0.7,
-        test_percentage: float=0.2) -> None:
+        train_percentage: float = 0.7,
+        test_percentage: float = 0.2) -> None:
+    """
+    Creates train, test, and validation sets from the given input directory and outputs
+    them to the specified output directory.
+
+    Parameters:
+        input_dir (Path): The directory containing the input images organized in
+        subdirectories per class.
+        output_dir (Path): The directory where the train, test, and validation sets
+        will be saved.
+        file_filter (str): The glob pattern to filter files in the input directory.
+        train_percentage (float): The percentage of images to be used for the
+        training set.
+        test_percentage (float): The percentage of images to be used for the
+        testing set.
+
+    Returns:
+        None
+    """
     output_folders = [_create_output_folder(output_dir, folder_name)
                       for folder_name in [TRAIN_FOLDER, TEST_FOLDER, VAL_FOLDER]]
 
@@ -33,20 +73,22 @@ def create_train_test_validation_sets(
         _copy_files_to_output(output_folders[1], image_class_folder, test_files)
         _copy_files_to_output(output_folders[2], image_class_folder, validation_files)
 
+
[…]

### 2024-09-19 project/utils/image_utils.py
Ensure valid percentage inputs for file splitting
Implemented checks to validate train and test percentages in the `create_train_test_validation_sets` function. Added corresponding unit tests to verify that invalid percentage inputs raise appropriate errors.

@@ -38,23 +38,34 @@ def create_train_test_validation_sets(
         train_percentage: float = 0.7,
         test_percentage: float = 0.2) -> None:
     """
-    Creates train, test, and validation sets from the given input directory and outputs
-    them to the specified output directory.
+        Splits and copies files from an input directory into train, test, and
+        validation sets  based on given percentages, and stores them in an output
+        directory.
 
-    Parameters:
-        input_dir (Path): The directory containing the input images organized in
-        subdirectories per class.
-        output_dir (Path): The directory where the train, test, and validation sets
-        will be saved.
-        file_filter (str): The glob pattern to filter files in the input directory.
-        train_percentage (float): The percentage of images to be used for the
-        training set.
-        test_percentage (float): The percentage of images to be used for the
-        testing set.
+        Parameters:
+        input_dir (Path): Path to the input directory containing class sub-folders with
+        files.
+        output_dir (Path): Path to the output directory where train, test, and
+        validation sub-folders will be created.
+        file_filter (str): Glob pattern to filter files in the input directory.
+        Default is "*.jpeg".
+        train_percentage (float): Percentage of files to be used for training.
+        Default is 0.7.
+        test_percentage (float): Percentage of files to be used for testing.
+        Default is 0.2.
 
-    Returns:
-        None
+        Raises:
+        ValueError: If train_percentage or test_percentage are not between 0 and 1.
     """
+
+    _ensure_valid_pct(
+        train_percentage,
+        "Training percentage must be greater than 0 and less than 1.")
+
+    _ensure_valid_pct(
+        test_percentage,
+        "Test percentage must be greater than 0 and less than 1.")
+
     output_folders = [_create_output_folder(output_dir, folder_name)
                       for folder_name in [TRAIN_FOLDER, TEST_FOLDER, VAL_FOLDER]]
 
@@ -72,6 +83,11 @@ def create_train_test_validation_sets(
         _copy_files_to_output(output_folders[0], image_class_folder, train_files)
         _copy_files_to_output(output_folders[1], image_class_folder, test_files)
         _copy_files_to_output(output_folders[2], image_class_folder, validation_files)
+
+
+def _ensure_valid_pct(pct:float, msg:str):
+    if not 0 < pct < 1:
+        raise ValueError(msg)
 
 
 def _copy_files_to_output(

### 2024-09-19 tests/test_image_utils.py

@@ -3,6 +3,7 @@ import shutil
 import unittest
 from pathlib import Path
 
+import pytest
 from parameterized import parameterized
 
 from project.utils.image_utils import create_train_test_validation_sets
@@ -19,6 +20,9 @@ class TestImageUtils(unittest.TestCase):
         self._output_dir = Path("test_create_train_test_validation_sets")
 
     def tearDown(self):
+        if not os.path.exists(self._output_dir):
+            return
+
         shutil.rmtree(self._output_dir)
 
     def create_output_dir_if_not_exists(self):
@@ -46,6 +50,28 @@ class TestImageUtils(unittest.TestCase):
         assert self._get_file_count(train_folder) == expected_train_file_count
         assert self._get_file_count(test_folder) == expected_test_file_count
         assert self._get_file_count(val_folder) == expected_validation_file_count
+
+    @parameterized.expand([1, 1.1, 0, -0.01, -1])
+    def test_invalid_train_pct_raises_error(
+            self,
+            train_pct):
+
+        with pytest.raises(ValueError):
+            create_train_test_validation_sets(
+                self._input_dir,
+                self._output_dir,
+                train_percentage=train_pct)
+
+    @parameterized.expand([1, 1.1, 0, -0.01, -1])
+    def test_invalid_test_pct_raises_error(
+            self,
+            test_pct):
+
+        with pytest.raises(ValueError):
+            create_train_test_validation_sets(
+                self._input_dir,
+                self._output_dir,
+                test_percentage=test_pct)
 
     def _get_file_count(self, folder:Path) -> int:
         return len(list(folder.rglob(JPEG_FILTER)))

### 2024-09-19 project/utils/image_utils.py
Add validation for train and test percentages sum
Ensure the sum of train and test percentages does not exceed 1. This change includes raising a ValueError when the sum is invalid and adds corresponding unit tests to verify the validation.

@@ -55,7 +55,8 @@ def create_train_test_validation_sets(
         Default is 0.2.
 
         Raises:
-        ValueError: If train_percentage or test_percentage are not between 0 and 1.
+        ValueError: If train_percentage or test_percentage == 0 or 1
+        ValueError: If train_percentage + test_percentage is >= 1
     """
 
     _ensure_valid_pct(
@@ -65,6 +66,10 @@ def create_train_test_validation_sets(
     _ensure_valid_pct(
         test_percentage,
         "Test percentage must be greater than 0 and less than 1.")
+
+    if train_percentage + test_percentage >= 1:
+        msg = "Sum of training and test percentages cannot exceed 1."
+        raise ValueError(msg)
 
     output_folders = [_create_output_folder(output_dir, folder_name)
                       for folder_name in [TRAIN_FOLDER, TEST_FOLDER, VAL_FOLDER]]

### 2024-09-19 tests/test_image_utils.py

@@ -73,6 +73,18 @@ class TestImageUtils(unittest.TestCase):
                 self._output_dir,
                 test_percentage=test_pct)
 
+    @parameterized.expand([
+        (0.7, 0.3),
+        (0.6, 0.5),
+    ])
+    def test_train_test_pct_should_be_less_than_1(self, train_pct, test_pct):
+        with pytest.raises(ValueError):
+            create_train_test_validation_sets(
+                self._input_dir,
+                self._output_dir,
+                train_percentage=train_pct,
+                test_percentage=test_pct)
+
     def _get_file_count(self, folder:Path) -> int:
         return len(list(folder.rglob(JPEG_FILTER)))
 

### 2024-09-19 tests/test_image_utils.py
Remove unused constants from test_image_utils.py
Deleted redundant TRAIN_FILE_COUNT, TEST_FILE_COUNT, and VAL_FILE_COUNT constants from the test file. This cleanup improves maintainability and avoids potential confusion.

@@ -8,9 +8,6 @@ from parameterized import parameterized
 
 from project.utils.image_utils import create_train_test_validation_sets
 
-TRAIN_FILE_COUNT = 14
-TEST_FILE_COUNT = 4
-VAL_FILE_COUNT = 2
 JPEG_FILTER = "*.jpeg"
 
 

### 2024-09-19 project/utils/image_utils.py
Refactor: consolidate percentage validity checks
Merged separate percentage validation checks into a single function. This reduces redundancy and improves code maintainability by centralizing all related validations.

@@ -59,17 +59,7 @@ def create_train_test_validation_sets(
         ValueError: If train_percentage + test_percentage is >= 1
     """
 
-    _ensure_valid_pct(
-        train_percentage,
-        "Training percentage must be greater than 0 and less than 1.")
-
-    _ensure_valid_pct(
-        test_percentage,
-        "Test percentage must be greater than 0 and less than 1.")
-
-    if train_percentage + test_percentage >= 1:
-        msg = "Sum of training and test percentages cannot exceed 1."
-        raise ValueError(msg)
+    _ensure_valid_pct(train_percentage, test_percentage)
 
     output_folders = [_create_output_folder(output_dir, folder_name)
                       for folder_name in [TRAIN_FOLDER, TEST_FOLDER, VAL_FOLDER]]
@@ -90,8 +80,17 @@ def create_train_test_validation_sets(
         _copy_files_to_output(output_folders[2], image_class_folder, validation_files)
 
 
-def _ensure_valid_pct(pct:float, msg:str):
-    if not 0 < pct < 1:
+def _ensure_valid_pct(train_percentage:float, test_percentage:float):
+    if not 0 < train_percentage < 1:
+        msg = "Training percentage must be greater than 0 and less than 1."
+        raise ValueError(msg)
+
+    if not 0 < test_percentage < 1:
+        msg = "Test percentage must be greater than 0 and less than 1."
+        raise ValueError(msg)
+
+    if train_percentage + test_percentage >= 1:
+        msg = "Sum of training and test percentages cannot exceed 1."
         raise ValueError(msg)
 
 

### 2024-09-19 project/utils/image_utils.py
Rename validation function for clarity
Renamed the `_ensure_valid_pct` function to `_ensure_valid_percentages` for better readability and clear identification of its purpose. This improves code maintainability by making function names more descriptive and understandable.

@@ -59,7 +59,7 @@ def create_train_test_validation_sets(
         ValueError: If train_percentage + test_percentage is >= 1
     """
 
-    _ensure_valid_pct(train_percentage, test_percentage)
+    _ensure_valid_percentages(train_percentage, test_percentage)
 
     output_folders = [_create_output_folder(output_dir, folder_name)
                       for folder_name in [TRAIN_FOLDER, TEST_FOLDER, VAL_FOLDER]]
@@ -80,7 +80,7 @@ def create_train_test_validation_sets(
         _copy_files_to_output(output_folders[2], image_class_folder, validation_files)
 
 
-def _ensure_valid_pct(train_percentage:float, test_percentage:float):
+def _ensure_valid_percentages(train_percentage:float, test_percentage:float):
     if not 0 < train_percentage < 1:
         msg = "Training percentage must be greater than 0 and less than 1."
         raise ValueError(msg)

### 2024-09-20 project/utils/image_utils.py
Rename function for better clarity.
Renamed `create_train_test_validation_sets` to `split_and_organize_images` to clearly reflect its purpose of splitting and organizing images. Updated corresponding function references in `image_utils.py` and its test file.

@@ -31,7 +31,7 @@ VAL_FOLDER = "val"
 SHUFFLE_ROUNDS = 3
 
 
-def create_train_test_validation_sets(
+def split_and_organize_images(
         input_dir: Path,
         output_dir: Path,
         file_filter="*.jpeg",

### 2024-09-20 tests/test_image_utils.py

@@ -6,7 +6,7 @@ from pathlib import Path
 import pytest
 from parameterized import parameterized
 
-from project.utils.image_utils import create_train_test_validation_sets
+from project.utils.image_utils import split_and_organize_images
 
 JPEG_FILTER = "*.jpeg"
 
@@ -38,7 +38,7 @@ class TestImageUtils(unittest.TestCase):
             expected_validation_file_count):
         self.create_output_dir_if_not_exists()
 
-        create_train_test_validation_sets(self._input_dir, self._output_dir)
+        split_and_organize_images(self._input_dir, self._output_dir)
 
         train_folder = Path(os.path.join(self._output_dir, "train", image_class))
         test_folder = Path(os.path.join(self._output_dir, "test", image_class))
@@ -54,7 +54,7 @@ class TestImageUtils(unittest.TestCase):
             train_pct):
 
         with pytest.raises(ValueError):
-            create_train_test_validation_sets(
+            split_and_organize_images(
                 self._input_dir,
                 self._output_dir,
                 train_percentage=train_pct)
@@ -65,7 +65,7 @@ class TestImageUtils(unittest.TestCase):
             test_pct):
 
         with pytest.raises(ValueError):
-            create_train_test_validation_sets(
+            split_and_organize_images(
                 self._input_dir,
                 self._output_dir,
                 test_percentage=test_pct)
@@ -76,7 +76,7 @@ class TestImageUtils(unittest.TestCase):
     ])
     def test_train_test_pct_should_be_less_than_1(self, train_pct, test_pct):
         with pytest.raises(ValueError):
-            create_train_test_validation_sets(
+            split_and_organize_images(
                 self._input_dir,
                 self._output_dir,
                 train_percentage=train_pct,

### 2024-09-20 project/utils/image_utils.py
Add function to generate image data with augmentation
Introduced `create_data_set_from_images` function in `image_utils.py` for generating batches of tensor image data with real-time data augmentation. This leverages TensorFlow's ImageDataGenerator and supports customizable target size, batch size, and class mode.

@@ -25,6 +25,8 @@ import shutil
 from pathlib import Path
 from typing import List, Tuple
 
+from tensorflow.keras.preprocessing.image import ImageDataGenerator
+
 TRAIN_FOLDER = "train"
 TEST_FOLDER = "test"
 VAL_FOLDER = "val"
@@ -78,6 +80,39 @@ def split_and_organize_images(
         _copy_files_to_output(output_folders[0], image_class_folder, train_files)
         _copy_files_to_output(output_folders[1], image_class_folder, test_files)
         _copy_files_to_output(output_folders[2], image_class_folder, validation_files)
+
+def create_data_set_from_images(
+        input_dir: Path,
+        target_size: Tuple[int, int]=(224,224),
+        batch_size:int=32,
+        class_mode:str="binary"):
+    """
+        Generates batches of tensor image data from the images in the specified
+        directory, with real-time data augmentation.
+
+        Parameters:
+        input_dir (Path): Path to the directory containing the images.
+        target_size (Tuple[int, int], optional): Tuple specifying the target size of
+        the images after resizing. Defaults to (224, 224).
+        batch_size (int, optional): Number of images to be yielded from the generator
+        per batch. Defaults to 32.
+        class_mode (str, optional): Mode for yielding the targets. Defaults to "binary".
+
+        Returns:
+        DirectoryIterator: An instance of DirectoryIterator yielding tuples of
+        (image_batch, label_batch).
+    """
+    image_data_gen = ImageDataGenerator(
+        rescale=1. / 255,
+        shear_range=0.2,
+        zoom_range=0.2,
+        horizontal_flip=True)
+
+    return image_data_gen.flow_from_directory(
+        input_dir,
+        target_size=target_size,
+        batch_size=batch_size,
+        class_mode=class_mode)
 
 
 def _ensure_valid_percentages(train_percentage:float, test_percentage:float):

### 2024-09-20 project/utils/image_utils.py
Add multithreading support to image splitting function
Integrated `concurrent.futures.ThreadPoolExecutor` to `split_and_organize_images` for parallel processing with configurable worker threads. Added `max_workers` parameter and corresponding validation to ensure it is greater than or equal to 1. Included unit tests to verify behavior with invalid worker counts.

@@ -24,6 +24,7 @@ import random
 import shutil
 from pathlib import Path
 from typing import List, Tuple
+from concurrent.futures import ThreadPoolExecutor
 
 from tensorflow.keras.preprocessing.image import ImageDataGenerator
 
@@ -38,48 +39,53 @@ def split_and_organize_images(
         output_dir: Path,
         file_filter="*.jpeg",
         train_percentage: float = 0.7,
-        test_percentage: float = 0.2) -> None:
+        test_percentage: float = 0.2,
+        max_workers:int=1) -> None:
     """
-        Splits and copies files from an input directory into train, test, and
-        validation sets  based on given percentages, and stores them in an output
-        directory.
+    Splits and copies files from an input directory into train, test, and
+    validation sets based on given percentages, and stores them in an output
+    directory.
 
-        Parameters:
-        input_dir (Path): Path to the input directory containing class sub-folders with
-        files.
-        output_dir (Path): Path to the output directory where train, test, and
-        validation sub-folders will be created.
-        file_filter (str): Glob pattern to filter files in the input directory.
-        Default is "*.jpeg".
-        train_percentage (float): Percentage of files to be used for training.
-        Default is 0.7.
-        test_percentage (float): Percentage of files to be used for testing.
-        Default is 0.2.
+    Parameters:
+    input_dir (Path): Path to the input directory containing class sub-folders with files.
+    output_dir (Path): Path to the output directory where train, test, and
+    validation sub-folders will be created.
+    file_filter (str): Glob pattern to filter files in the input directory.
+    Default is "*.jpeg".
+    train_percentage (float): Percentage of files to be used for training.
+    Default is 0.7.
+    test_percentage (float): Percentage of files to be used for testing.
+    Default is 0.2.
+    max_workers (int): Number of worker threads to use. Default is 1.
 
-        Raises:
-        ValueError: If train_percentage or test_percentage == 0 or 1
-        ValueError: If train_percentage + test_percentage is >= 1
+    Raises:
+    ValueError: If train_percentage or test_percentage == 0 or 1
+    ValueError: If train_percentage + test_percentage is >= 1
     """
-
     _ensure_valid_percentages(train_percentage, test_percentage)
+
+    if max_workers < 1:
+        raise ValueError("max_workers must be greater than or equal to 1.")
 
     output_folders = [_create_output_folder(output_dir, folder_name)
[…]

### 2024-09-20 tests/test_image_utils.py

@@ -70,6 +70,14 @@ class TestImageUtils(unittest.TestCase):
                 self._output_dir,
                 test_percentage=test_pct)
 
+    @parameterized.expand([0, -1])
+    def test_invalid_worker_raises_error(self, max_workers):
+        with pytest.raises(ValueError):
+            split_and_organize_images(
+                self._input_dir,
+                self._output_dir,
+                max_workers=max_workers)
+
     @parameterized.expand([
         (0.7, 0.3),
         (0.6, 0.5),

### 2024-09-21 project/utils/image_utils.py
Change class_mode to "categorical" in image generator
Updated the class_mode parameter in the image data generator function from "binary" to "categorical". This allows for multi-class classification, enhancing the flexibility of the data preprocessing steps.

@@ -91,7 +91,7 @@ def create_data_set_from_images(
         input_dir: Path,
         target_size: Tuple[int, int]=(224,224),
         batch_size:int=32,
-        class_mode:str="binary"):
+        class_mode:str="categorical"):
     """
         Generates batches of tensor image data from the images in the specified
         directory, with real-time data augmentation.

### 2024-09-21 project/utils/cnn_builder.py
Added progress reports.

@@ -1,63 +1,0 @@
-from tensorflow.keras.models import Sequential
-from tensorflow.keras.layers import Conv2D, MaxPool2D, Flatten, Dense, Dropout
-from typing import Tuple, List
-
-class CNNBuilder:
-    def __init__(self):
-        self._model = Sequential()
-
-    def add_input(
-            self,
-            filters:int=64,
-            kernel_size:int=3,
-            strides:Tuple[int, int]=(2, 2),
-            activation:str="relu",
-            input_shape:List[int]=[224,224,3]) -> None:
-
-        self._model.add(Conv2D(
-            filters=filters,
-            kernel_size=kernel_size,
-            strides=strides,
-            activation=activation,
-            input_shape=input_shape))
-
-    def add_max_pool(
-            self,
-            pool_size:Tuple[int,int]=(3, 3),
-            strides:int=2,
-            padding:str="same") -> None:
-
-        self._model.add(MaxPool2D(
-            pool_size=pool_size, strides=strides, padding=padding))
-
-    def add_convolution(
-            self,
-            filters: int = 64,
-            kernel_size: int = 3,
-            strides: Tuple[int, int] = (2, 2),
-            activation: str = "relu",
-            padding:str="same") -> None:
-
-        self._model.add(Conv2D(
-            filters=filters,
-            kernel_size=kernel_size,
-            strides=strides,
-            activation=activation,
-            padding=padding))
-
-    def add_flatten(self):
-        self._model.add(Flatten())
-
-    def add_dropout(self, rate:float):
-         self._model.add(Dropout(rate))
-
-    def add_output_layer(self, num_classes:int, activation:str="relu"):
-        self._model.add(Dense(units=num_classes, activation=activation))
-
-    def build(self):
-        self._model.compile(
-            optimizer='adam',
-            loss='binary_crossentropy',
-            metrics=['accuracy'])
-
-        return self._model

### 2024-09-21 project/utils/download_synsets.py

@@ -1,153 +1,0 @@
-"""
-Module: download_synsets
-
-This module provides functionality to download and rename synset tar files from
-ImageNet.
-It utilizes asynchronous operations to handle multiple downloads concurrently, making
-the download process efficient and quicker.
-
-Logging is configured to display info level messages which will help in tracking the
-progress of downloads.
-
-The module defines a main function which orchestrates the download process for all the
-synsets specified in the synset_mapping dictionary. Each synset is downloaded and
-renamed according to its CIFAR-10 class name.
-
-The primary components of this module include:
-
-- ROOT_URL: The base URL from where the ImageNet tar files are downloaded.
-- synset_mapping: A dictionary mapping CIFAR-10 class names to their corresponding
-  synset identifiers in ImageNet.
-- download_dir: The directory where the downloaded tar files will be stored.
-
-Functions:
-- download_and_rename_tar: This function handles the actual downloading and renaming
-  of a synset tar file using aiohttp for asynchronous HTTP requests.
-- main: The main function that initializes the aiohttp ClientSession and triggers
-  the download of all synsets concurrently.
-
-Usage:
-Run the module as a standalone script to initiate the download process:
-    $ python download_synsets.py
-
-Dependencies:
-- aiohttp: Asynchronous HTTP client/server framework
-- asyncio: Python standard library module for writing asynchronous code
-- logging: Python standard library module for logging messages
-- os: Python standard library module for operating system-related functionalities
-"""
-import argparse
-import asyncio
-import logging
-import os
-
-import aiohttp
-
-logging.basicConfig(level=logging.INFO,
-                    format="%(asctime)s - %(levelname)s - %(message)s")
-
-ROOT_URL = "https://image-net.org/data/winter21_whole/"
-
-synset_mapping = {
-    "airplane": "n02691156",
-    "automobile": "n02958343",
-    "bird": "n01503061",
-    "cat": "n02121808",
-    "deer": "n02419796",
-    "dog": "n02084071",
-    "frog": "n01641577",
-    "horse": "n02374451",
-    "ship": "n04194289",
-    "truck": "n04467665",
-}
-
-async def download_and_rename_tar(session, synset_id, cifar_class_name, download_dir):
-    """
-    Asynchronously downloads a .tar file from a given URL, renames the file,
-    and saves it to a specified directory.
-
-    Parameters:
-    session (aiohttp.ClientSession): The aiohttp session used for making HTTP requests.
-    synset_id (str): The synset ID used to construct the download URL.
[…]

### 2024-09-21 project/utils/preprocess_images.py

@@ -1,230 +1,0 @@
-"""
-Module for preprocessing images.
-
-This module includes functions to resize and find images, as well as a main function
-to process images as a whole.
-
-Functions:
-    - resize_image: Resizes an image to given dimensions.
-    - find_images: Identifies and retrieves images from a specified directory.
-    - process_images: Handles the overall processing workflow for images.
-    - main: Entry point for the module, coordinating the complete image
-    preprocessing pipeline.
-"""
-import argparse
-import logging
-import time
-from collections import defaultdict
-from concurrent.futures import ProcessPoolExecutor
-from pathlib import Path
-from typing import Tuple
-
-from PIL import Image
-
-
-def setup_logging():
-    """
-
-        Configures and sets up logging for the image resizer application.
-
-        The configuration includes:
-        - Creating a logger named 'image_resizer' with a logging level of INFO.
-        - Creating a console handler for outputting logs to the console.
-        - Creating a file handler for writing logs to a file named 'image_resize.log'.
-        - Setting a log message format that includes the timestamp, logger name, log
-        level, and the actual message.
-        - Adding both the console handler and file handler to the logger if they haven't
-        been added already.
-
-        Returns:
-            console_handler: The configured console handler.
-            file_handler: The configured file handler.
-            logger: The logger configured for the image resizer application.
-
-    """
-    logger = logging.getLogger("image_resizer")
-    logger.setLevel(logging.INFO)
-
-    # Create handlers
-    console_handler = logging.StreamHandler()
-    file_handler = logging.FileHandler("image_resize.log")
-
-    # Create formatters and add them to handlers
-    formatter = logging.Formatter(
-        "%(asctime)s - %(name)s - %(levelname)s - %(message)s")
-    console_handler.setFormatter(formatter)
-    file_handler.setFormatter(formatter)
-
-    # Add handlers to the logger
-    if not logger.hasHandlers():
-        logger.addHandler(console_handler)
-        logger.addHandler(file_handler)
-
-    return console_handler, file_handler, logger
-
-def resize_image(image_path:Path, output_path:Path, size:Tuple[int,int]):
-    """
-    resize_image(image_path, output_path, size)
-
-    Resizes the image located at the given image path and saves the resized image to
-    the specified output path.
-
-    Parameters:
-    image_path (Path): The file path to the original image.
-    output_path (Path): The file path to save the resized image.
[…]

### 2024-09-21 tests/test_cnn_builder.py

@@ -1,27 +1,0 @@
-import sys
-import unittest
-from sys import modules
-from unittest.mock import MagicMock, patch
-
-from project.utils.cnn_builder import CNNBuilder
-
-
-class TestCNNBuilder(unittest.TestCase):
-
-    def setUp(self):
-        self.builder = CNNBuilder()
-
-    def test_build(self):
-        self.builder.add_input(kernel_size=7)
-        self.builder.add_max_pool()
-        self.builder.add_convolution(filters=192, kernel_size=3, strides=(1, 1))
-        self.builder.add_max_pool()
-        self.builder.add_dropout(0.4)
-        self.builder.add_flatten()
-        self.builder.add_output_layer(10, activation="softmax")
-
-        model = self.builder.build()
-        assert len(model.layers) == 7
-
-if __name__ == "__main__":
-    unittest.main()

### 2024-09-21 tests/test_preprocess_images.py

@@ -1,73 +1,0 @@
-import os
-import unittest
-from pathlib import Path
-from unittest.mock import MagicMock
-
-import pytest
-from parameterized import parameterized
-from PIL import Image
-
-from project.utils import preprocess_images
-
-
-class TestPreprocessImages(unittest.TestCase):
-    def setUp(self):
-        self.log_mock = MagicMock()
-
-    def tearDown(self):
-        self.log_mock.reset_mock()
-
-    def create_dummy_image(self, path, size):
-        with Image.new("RGB", size) as img:
-            img.save(path)
-
-    @parameterized.expand([
-        ((32, 32),),
-        ((224, 224),),
-        ((512, 32),),
-        ((32, 512),),
-    ])
-    def test_resize_image(self, original_size):
-        image_path = Path("test_original.jpg")
-        self.create_dummy_image(image_path, original_size)
-
-        output_path = Path("test_resized.jpg")
-        output_size = (224, 224)
-        preprocess_images.resize_image(image_path, output_path, output_size)
-
-        with Image.open(output_path) as img:
-            assert img.size == output_size
-
-        os.remove(image_path)
-        os.remove(output_path)
-
-    @parameterized.expand([
-        (0,),
-        (-1,),
-    ])
-    def test_invalid_workers(self, workers):
-        with pytest.raises(ValueError):
-            preprocess_images.process_images(
-                Path("inputdir"),
-                Path("outputdir"),
-                (10, 10),
-                self.log_mock,
-                workers)
-
-    @parameterized.expand([
-        ((0, 0),),
-        ((-1, 1),),
-        ((1, -1),),
-    ])
-    def test_invalid_size(self, size):
-        with pytest.raises(ValueError):
-            preprocess_images.process_images(
-                Path("inputdir"),
-                Path("outputdir"),
-                size,
-                self.log_mock)
-
-
-
-if __name__ == "__main__":
-    unittest.main()

### 2024-09-21 project/progress_summary.tex
Added progress report and in-progress report.

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{cite}

\graphicspath{{project/paper_images/}}

\title{Deep Learning Project - Progress Summary}
\author{
    Group Members: \\
    Farooq Mahmud
}
\date{September 20, 2024}

\begin{document}

\maketitle

\section{Progress summary}
\begin{itemize}
    \item Read papers cited in working version of report.
    \item Downloaded subset of Imagenet images that mapped to CIFAR-10 image classes. 
    \item Scaled images to 224 x 244 which matches the input size of the Szegedy paper.
    \item Trained a simple Sequential model using downloaded images.
    \item Reading Tensorflow documentation on how to create the more complex model in the Szegedy paper.
\end{itemize}

\end{document}

### 2024-09-21 project/paper.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{cite}

\graphicspath{{project/paper_images/}}

\title{Deep Learning Project}
\author{
    Group Members: \\
    Farooq Mahmud
}
\date{September 20, 2024}

\begin{document}

\maketitle

\section{Introduction}
Convolutional Neural Networks (CNNs) have become a cornerstone in the field of computer vision, enabling significant advancements in tasks such as image classification. One of the most impactful architectures is the **Inception Network** introduced in the paper ``Going Deeper with Convolutions" by Szegedy et al.~\cite{szegedy2015going}. This architecture, commonly referred to as **GoogLeNet**, achieved remarkable performance in the 2014 ImageNet competition by introducing the concept of multi-scale processing through Inception modules.

The objective of this project is to analyze and replicate the Inception v1 (GoogLeNet) model using TensorFlow, train it on a subset of the ImageNet dataset, and evaluate its performance. Additionally, we will adapt the model to other datasets, such as CIFAR-10, to test its generalization capabilities.

\section{Background}

GoogLeNet introduced a novel approach to convolutional neural networks by stacking multiple Inception modules. Each Inception module consists of parallel branches that apply different-sized convolutional filters (1x1, 3x3, 5x5) and max-pooling operations in parallel. These branches are concatenated to capture multi-scale features.

The Inception module aims to mitigate computational costs by including 1x1 convolutions for dimensionality reduction before performing the larger convolution operations. This was one of the key innovations that made GoogLeNet both efficient and powerful, despite being deeper than previous CNN architectures such as AlexNet and VGGNet~\cite{krizhevsky2012imagenet, simonyan2015vgg}.

\section{Methodology}

\subsection{Image Selection}
The Imagenet dataset contains over 1 million images with a size of over 1TB. Processing this amount of data would not be practical given the available coumputing power. In order to make processing possible in a reasonable timeframe, a subset of the Imagenet images was downloaded. The subsets, or \textit{synsets} map to the CIFAR-10 image classes as shown in Table \ref{tab:synsets}.

\begin{table}[ht]
    \centering
    \begin{tabular}{ll}
    \hline
    \textbf{Synset ID} & \textbf{Equivalent CIFAR-10 Class} \\ \hline
    n02691156 & airplane \\ \hline
    n02958343 & automobile \\ \hline
    n01503061 & bird \\ \hline
    n02121808 & cat \\ \hline
    n02419796 & deer \\ \hline
    n02084071 & dog \\ \hline
    n01641577 & frog \\ \hline
    n02374451 & horse \\ \hline
    n04194289 & ship \\ \hline
    n04467665 & truck \\ \hline
    \end{tabular}
    \caption{Imagenet synset to CIFAR-10 class mapping}
    \label{tab:synsets}
\end{table}

The resulting download consists of 14,738 files with a total size of 1.3GB. Scaling the images to 224 x 224 pixels, the dimensions used by Szegedy, et. al. decreased the size to a manageable 137MB.

After scaling, the images were organized into training, testing, and validation sets. The percentages were 70\% for training, 20\% for testing, and 10\% for validation.

\subsection{A Simple Sequential Network as a Control}
Before constructing a complex model like GoogLeNet, it is a good idea to use a simple network as a baseline. This network is shown in Figure \ref{fig:cnn-control}

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.7]{project/paper_images/control_cnn.png}
    \caption{Simple CNN used as a control}
    \label{fig:cnn-control}
\end{figure}

This simple model, when trained wfor 25 epochs, achieved a maximum accuracy of 59\% as shown in Figure \ref{fig:cnn-control-accuracy}.

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.7]{project/paper_images/control_cnn_accuracy.png}
    \caption{Control CNN accuracy}
    \label{fig:cnn-control-accuracy}
\end{figure}

\clearpage

\subsection{Replicating the Inception v1 (GoogLeNet) Model}
The Inception v1 model incorporates branching and parallel operations. As a result, Tensorflow's Sequential  model is inappropriate. Tensorflow's functional API must be used instead.

\bibliographystyle{plain}
\bibliography{project/refs}

\end{document}

### 2024-09-21 project/utils/cnn_builder.py
Adding back deleted files.

@@ -1,0 +1,63 @@
+from tensorflow.keras.models import Sequential
+from tensorflow.keras.layers import Conv2D, MaxPool2D, Flatten, Dense, Dropout
+from typing import Tuple, List
+
+class CNNBuilder:
+    def __init__(self):
+        self._model = Sequential()
+
+    def add_input(
+            self,
+            filters:int=64,
+            kernel_size:int=3,
+            strides:Tuple[int, int]=(2, 2),
+            activation:str="relu",
+            input_shape:List[int]=[224,224,3]) -> None:
+
+        self._model.add(Conv2D(
+            filters=filters,
+            kernel_size=kernel_size,
+            strides=strides,
+            activation=activation,
+            input_shape=input_shape))
+
+    def add_max_pool(
+            self,
+            pool_size:Tuple[int,int]=(3, 3),
+            strides:int=2,
+            padding:str="same") -> None:
+
+        self._model.add(MaxPool2D(
+            pool_size=pool_size, strides=strides, padding=padding))
+
+    def add_convolution(
+            self,
+            filters: int = 64,
+            kernel_size: int = 3,
+            strides: Tuple[int, int] = (2, 2),
+            activation: str = "relu",
+            padding:str="same") -> None:
+
+        self._model.add(Conv2D(
+            filters=filters,
+            kernel_size=kernel_size,
+            strides=strides,
+            activation=activation,
+            padding=padding))
+
+    def add_flatten(self):
+        self._model.add(Flatten())
+
+    def add_dropout(self, rate:float):
+         self._model.add(Dropout(rate))
+
+    def add_output_layer(self, num_classes:int, activation:str="relu"):
+        self._model.add(Dense(units=num_classes, activation=activation))
+
+    def build(self):
+        self._model.compile(
+            optimizer='adam',
+            loss='binary_crossentropy',
+            metrics=['accuracy'])
+
+        return self._model

### 2024-09-21 project/utils/download_synsets.py

@@ -1,0 +1,153 @@
+"""
+Module: download_synsets
+
+This module provides functionality to download and rename synset tar files from
+ImageNet.
+It utilizes asynchronous operations to handle multiple downloads concurrently, making
+the download process efficient and quicker.
+
+Logging is configured to display info level messages which will help in tracking the
+progress of downloads.
+
+The module defines a main function which orchestrates the download process for all the
+synsets specified in the synset_mapping dictionary. Each synset is downloaded and
+renamed according to its CIFAR-10 class name.
+
+The primary components of this module include:
+
+- ROOT_URL: The base URL from where the ImageNet tar files are downloaded.
+- synset_mapping: A dictionary mapping CIFAR-10 class names to their corresponding
+  synset identifiers in ImageNet.
+- download_dir: The directory where the downloaded tar files will be stored.
+
+Functions:
+- download_and_rename_tar: This function handles the actual downloading and renaming
+  of a synset tar file using aiohttp for asynchronous HTTP requests.
+- main: The main function that initializes the aiohttp ClientSession and triggers
+  the download of all synsets concurrently.
+
+Usage:
+Run the module as a standalone script to initiate the download process:
+    $ python download_synsets.py
+
+Dependencies:
+- aiohttp: Asynchronous HTTP client/server framework
+- asyncio: Python standard library module for writing asynchronous code
+- logging: Python standard library module for logging messages
+- os: Python standard library module for operating system-related functionalities
+"""
+import argparse
+import asyncio
+import logging
+import os
+
+import aiohttp
+
+logging.basicConfig(level=logging.INFO,
+                    format="%(asctime)s - %(levelname)s - %(message)s")
+
+ROOT_URL = "https://image-net.org/data/winter21_whole/"
+
+synset_mapping = {
+    "airplane": "n02691156",
+    "automobile": "n02958343",
+    "bird": "n01503061",
+    "cat": "n02121808",
+    "deer": "n02419796",
+    "dog": "n02084071",
+    "frog": "n01641577",
+    "horse": "n02374451",
+    "ship": "n04194289",
+    "truck": "n04467665",
+}
+
+async def download_and_rename_tar(session, synset_id, cifar_class_name, download_dir):
+    """
+    Asynchronously downloads a .tar file from a given URL, renames the file,
+    and saves it to a specified directory.
+
+    Parameters:
+    session (aiohttp.ClientSession): The aiohttp session used for making HTTP requests.
+    synset_id (str): The synset ID used to construct the download URL.
[…]

### 2024-09-21 project/utils/image_utils.py

@@ -1,0 +1,192 @@
+"""
+Module: image_utils
+
+This module contains utilities for creating train, test, and validation sets
+from a directory of images. The sets are created based on given percentages
+for training and testing, with the remainder being used for validation.
+
+Functions:
+    create_train_test_validation_sets: Main function to create the dataset splits.
+    _copy_files_to_output: Helper function to copy files to the output directory.
+    _split_files: Helper function to split the files into train, test, and
+    validation sets.
+    _get_source_files: Helper function to retrieve source files from a directory.
+    _get_files_from_folder: Helper function to get files from a specific folder.
+    _copy_files: Helper function to copy files to the specified output folder.
+    _create_output_folder: Helper function to create an output directory if it
+    doesn't exist.
+    _create_image_class_output_folder: Helper function to create sub-directories within
+    the output directory.
+"""
+
+import os.path
+import random
+import shutil
+from pathlib import Path
+from typing import List, Tuple
+from concurrent.futures import ThreadPoolExecutor
+
+from tensorflow.keras.preprocessing.image import ImageDataGenerator
+
+TRAIN_FOLDER = "train"
+TEST_FOLDER = "test"
+VAL_FOLDER = "val"
+SHUFFLE_ROUNDS = 3
+
+
+def split_and_organize_images(
+        input_dir: Path,
+        output_dir: Path,
+        file_filter="*.jpeg",
+        train_percentage: float = 0.7,
+        test_percentage: float = 0.2,
+        max_workers:int=1) -> None:
+    """
+    Splits and copies files from an input directory into train, test, and
+    validation sets based on given percentages, and stores them in an output
+    directory.
+
+    Parameters:
+    input_dir (Path): Path to the input directory containing class sub-folders with files.
+    output_dir (Path): Path to the output directory where train, test, and
+    validation sub-folders will be created.
+    file_filter (str): Glob pattern to filter files in the input directory.
+    Default is "*.jpeg".
+    train_percentage (float): Percentage of files to be used for training.
+    Default is 0.7.
+    test_percentage (float): Percentage of files to be used for testing.
+    Default is 0.2.
+    max_workers (int): Number of worker threads to use. Default is 1.
+
+    Raises:
+    ValueError: If train_percentage or test_percentage == 0 or 1
+    ValueError: If train_percentage + test_percentage is >= 1
+    """
[…]

### 2024-09-21 project/utils/preprocess_images.py

@@ -1,0 +1,230 @@
+"""
+Module for preprocessing images.
+
+This module includes functions to resize and find images, as well as a main function
+to process images as a whole.
+
+Functions:
+    - resize_image: Resizes an image to given dimensions.
+    - find_images: Identifies and retrieves images from a specified directory.
+    - process_images: Handles the overall processing workflow for images.
+    - main: Entry point for the module, coordinating the complete image
+    preprocessing pipeline.
+"""
+import argparse
+import logging
+import time
+from collections import defaultdict
+from concurrent.futures import ProcessPoolExecutor
+from pathlib import Path
+from typing import Tuple
+
+from PIL import Image
+
+
+def setup_logging():
+    """
+
+        Configures and sets up logging for the image resizer application.
+
+        The configuration includes:
+        - Creating a logger named 'image_resizer' with a logging level of INFO.
+        - Creating a console handler for outputting logs to the console.
+        - Creating a file handler for writing logs to a file named 'image_resize.log'.
+        - Setting a log message format that includes the timestamp, logger name, log
+        level, and the actual message.
+        - Adding both the console handler and file handler to the logger if they haven't
+        been added already.
+
+        Returns:
+            console_handler: The configured console handler.
+            file_handler: The configured file handler.
+            logger: The logger configured for the image resizer application.
+
+    """
+    logger = logging.getLogger("image_resizer")
+    logger.setLevel(logging.INFO)
+
+    # Create handlers
+    console_handler = logging.StreamHandler()
+    file_handler = logging.FileHandler("image_resize.log")
+
+    # Create formatters and add them to handlers
+    formatter = logging.Formatter(
+        "%(asctime)s - %(name)s - %(levelname)s - %(message)s")
+    console_handler.setFormatter(formatter)
+    file_handler.setFormatter(formatter)
+
+    # Add handlers to the logger
+    if not logger.hasHandlers():
+        logger.addHandler(console_handler)
+        logger.addHandler(file_handler)
+
+    return console_handler, file_handler, logger
+
+def resize_image(image_path:Path, output_path:Path, size:Tuple[int,int]):
+    """
+    resize_image(image_path, output_path, size)
+
+    Resizes the image located at the given image path and saves the resized image to
+    the specified output path.
+
+    Parameters:
+    image_path (Path): The file path to the original image.
+    output_path (Path): The file path to save the resized image.
[…]

### 2024-09-21 tests/test_cnn_builder.py

@@ -1,0 +1,27 @@
+import sys
+import unittest
+from sys import modules
+from unittest.mock import MagicMock, patch
+
+from project.utils.cnn_builder import CNNBuilder
+
+
+class TestCNNBuilder(unittest.TestCase):
+
+    def setUp(self):
+        self.builder = CNNBuilder()
+
+    def test_build(self):
+        self.builder.add_input(kernel_size=7)
+        self.builder.add_max_pool()
+        self.builder.add_convolution(filters=192, kernel_size=3, strides=(1, 1))
+        self.builder.add_max_pool()
+        self.builder.add_dropout(0.4)
+        self.builder.add_flatten()
+        self.builder.add_output_layer(10, activation="softmax")
+
+        model = self.builder.build()
+        assert len(model.layers) == 7
+
+if __name__ == "__main__":
+    unittest.main()

### 2024-09-21 tests/test_image_utils.py

@@ -1,0 +1,97 @@
+import os
+import shutil
+import unittest
+from pathlib import Path
+
+import pytest
+from parameterized import parameterized
+
+from project.utils.image_utils import split_and_organize_images
+
+JPEG_FILTER = "*.jpeg"
+
+
+class TestImageUtils(unittest.TestCase):
+    def setUp(self):
+        self._input_dir = Path("./images224")
+        self._output_dir = Path("test_create_train_test_validation_sets")
+
+    def tearDown(self):
+        if not os.path.exists(self._output_dir):
+            return
+
+        shutil.rmtree(self._output_dir)
+
+    def create_output_dir_if_not_exists(self):
+        if not os.path.exists(self._output_dir):
+            os.makedirs(self._output_dir)
+
+    @parameterized.expand([
+        ("bird", 14, 4, 2),
+        ("cat", 14, 4, 2),
+    ])
+    def test_create_train_test_validation_sets(
+            self,
+            image_class,
+            expected_train_file_count,
+            expected_test_file_count,
+            expected_validation_file_count):
+        self.create_output_dir_if_not_exists()
+
+        split_and_organize_images(self._input_dir, self._output_dir)
+
+        train_folder = Path(os.path.join(self._output_dir, "train", image_class))
+        test_folder = Path(os.path.join(self._output_dir, "test", image_class))
+        val_folder = Path(os.path.join(self._output_dir, "val", image_class))
+
+        assert self._get_file_count(train_folder) == expected_train_file_count
+        assert self._get_file_count(test_folder) == expected_test_file_count
+        assert self._get_file_count(val_folder) == expected_validation_file_count
+
+    @parameterized.expand([1, 1.1, 0, -0.01, -1])
+    def test_invalid_train_pct_raises_error(
+            self,
+            train_pct):
+
+        with pytest.raises(ValueError):
+            split_and_organize_images(
+                self._input_dir,
+                self._output_dir,
+                train_percentage=train_pct)
+
+    @parameterized.expand([1, 1.1, 0, -0.01, -1])
+    def test_invalid_test_pct_raises_error(
+            self,
+            test_pct):
+
+        with pytest.raises(ValueError):
+            split_and_organize_images(
+                self._input_dir,
+                self._output_dir,
+                test_percentage=test_pct)
+
+    @parameterized.expand([0, -1])
+    def test_invalid_worker_raises_error(self, max_workers):
+        with pytest.raises(ValueError):
[…]

### 2024-09-21 tests/test_preprocess_images.py

@@ -1,0 +1,73 @@
+import os
+import unittest
+from pathlib import Path
+from unittest.mock import MagicMock
+
+import pytest
+from parameterized import parameterized
+from PIL import Image
+
+from project.utils import preprocess_images
+
+
+class TestPreprocessImages(unittest.TestCase):
+    def setUp(self):
+        self.log_mock = MagicMock()
+
+    def tearDown(self):
+        self.log_mock.reset_mock()
+
+    def create_dummy_image(self, path, size):
+        with Image.new("RGB", size) as img:
+            img.save(path)
+
+    @parameterized.expand([
+        ((32, 32),),
+        ((224, 224),),
+        ((512, 32),),
+        ((32, 512),),
+    ])
+    def test_resize_image(self, original_size):
+        image_path = Path("test_original.jpg")
+        self.create_dummy_image(image_path, original_size)
+
+        output_path = Path("test_resized.jpg")
+        output_size = (224, 224)
+        preprocess_images.resize_image(image_path, output_path, output_size)
+
+        with Image.open(output_path) as img:
+            assert img.size == output_size
+
+        os.remove(image_path)
+        os.remove(output_path)
+
+    @parameterized.expand([
+        (0,),
+        (-1,),
+    ])
+    def test_invalid_workers(self, workers):
+        with pytest.raises(ValueError):
+            preprocess_images.process_images(
+                Path("inputdir"),
+                Path("outputdir"),
+                (10, 10),
+                self.log_mock,
+                workers)
+
+    @parameterized.expand([
+        ((0, 0),),
+        ((-1, 1),),
+        ((1, -1),),
+    ])
+    def test_invalid_size(self, size):
+        with pytest.raises(ValueError):
+            preprocess_images.process_images(
+                Path("inputdir"),
+                Path("outputdir"),
+                size,
+                self.log_mock)
+
+
+
+if __name__ == "__main__":
+    unittest.main()

### 2024-09-23 project/utils/cnn_builder.py
Add method for adding fully connected layer to CNN builder
This commit introduces a new method `add_full_connection` to the CNN builder, allowing the addition of fully connected (Dense) layers with customizable units and activation functions. The Jupyter notebook is updated to include this new method, and minor adjustments are made to paths and data output formats to reflect the model changes.

@@ -51,6 +51,9 @@ class CNNBuilder:
     def add_dropout(self, rate:float):
          self._model.add(Dropout(rate))
 
+    def add_full_connection(self, units:int=128, activation:str="relu"):
+        self._model.add(Dense(units=units, activation=activation))
+
     def add_output_layer(self, num_classes:int, activation:str="relu"):
         self._model.add(Dense(units=num_classes, activation=activation))
 

### 2024-09-23 project/utils/download_synsets.py
Simplify renaming scheme in download function
Change the renaming logic to use only the CIFAR class name for the tar filename, instead of combining it with the synset ID. This improves clarity and removes redundancy from the file naming convention.

@@ -85,7 +85,7 @@ async def download_and_rename_tar(session, synset_id, cifar_class_name, download
     - Logs progress and any errors encountered during the download and renaming process.
     """
     tar_url = f"{ROOT_URL}{synset_id}.tar"
-    renamed_tar_filename = f"{synset_id}-{cifar_class_name}.tar"
+    renamed_tar_filename = f"{cifar_class_name}.tar"
     renamed_tar_file_path = os.path.join(download_dir, renamed_tar_filename)
 
     if os.path.exists(renamed_tar_file_path):

### 2024-09-23 project/utils/preprocess_images.py
Refactor preprocess_images.py for improved readability
Reorder function arguments to match the expected signature and convert input and output directories to Path objects. This ensures type consistency and improves

@@ -211,11 +211,11 @@ def main():
             f"{args.size[0]}x{args.size[1]}.")
 
         process_images(
-            args.input_dir,
-            args.output_dir,
+            Path(args.input_dir),
+            Path(args.output_dir),
             tuple(args.size),
-            args.workers,
-            logger)
+            logger,
+            args.workers)
 
         logger.info("Image resizing process completed.")
     except Exception:

### 2024-09-23 project/utils/cnn_builder.py
Add precision and recall metrics to CNN model
Updated the CNNBuilder class to include precision and recall as metrics when compiling the model to provide a more comprehensive evaluation. Also, modified the notebook to reflect these changes and provide more detailed training result outputs.

@@ -1,5 +1,6 @@
 from tensorflow.keras.models import Sequential
 from tensorflow.keras.layers import Conv2D, MaxPool2D, Flatten, Dense, Dropout
+from tensorflow.keras.metrics import Precision, Recall
 from typing import Tuple, List
 
 class CNNBuilder:
@@ -61,6 +62,6 @@ class CNNBuilder:
         self._model.compile(
             optimizer='adam',
             loss='binary_crossentropy',
-            metrics=['accuracy'])
+            metrics=["accuracy", Precision(name="precision"), Recall(name="recall")])
 
         return self._model

### 2024-09-28 project/paper.tex
Add new sections, revise text, and include Python code
Introduce sections on simplified and full GoogLeNet networks, refine objectives and background details, and embed code for augmentation. Simplify image set handling using TensorFlow and update the training methodology.

@@
 \usepackage{cite}
+\usepackage{listings}
+\usepackage{xcolor}
+
+\lstset{
+    language=Python,             % Specify Python as the language
+    basicstyle=\ttfamily,        % Use monospaced font for code
+    keywordstyle=\color{blue},   % Color for keywords
+    commentstyle=\color{gray},   % Color for comments
+    stringstyle=\color{red},     % Color for strings
+    numbers=left,                % Line numbers on the left
+    numberstyle=\tiny\color{gray}, % Style for line numbers
+    stepnumber=1,                % Number every line
+    frame=single,                % Frame around the code
+    breaklines=true              % Enable line breaking
+}
 
@@
 \section{Introduction}
-Convolutional Neural Networks (CNNs) have become a cornerstone in the field of computer vision, enabling significant advancements in tasks such as image classification.
-One of the most impactful architectures is the **Inception Network** introduced in the paper ``Going Deeper with Convolutions" by Szegedy et al.~\cite{szegedy2015going}.
-This architecture, commonly referred to as **GoogLeNet**, achieved remarkable performance in the 2014 ImageNet competition by introducing the concept of multi-scale processing through Inception modules.
+Deep learning models, particularly Convolutional Neural Networks, have made great strides in solving complex computer vision tasks such as image classification.
+Among the key advancements in this area is the Inception architecture, introduced by Szegedy et al.
+(2015) in the paper "Going Deeper with Convolutions." This architecture was designed to balance computational efficiency with improved classification accuracy.
+This architecture achieved winning results in the ImageNet Large-Scale Visual Recognition Challenge (ILSVRC) in 2014.
+The primary innovation of the Inception architecture is its use of multi-scale feature extraction and dimensionality reduction, which allows it to increase network depth and width without excessive computational overhead\cite{szegedy2015going}.
 
-The objective of this project is to analyze and replicate the Inception v1 (GoogLeNet) model using TensorFlow, train it on a subset of the ImageNet dataset, and evaluate its performance.
-Additionally, we will adapt the model to other datasets, such as CIFAR-10, to test its generalization capabilities.
+This paper aims to replicate the results presented by Szegedy et al., critically assess the impact of the Inception architecture, and test its performance on an alternative dataset.
+We attempt to reproduce the GoogLeNet model by writing our own Python implementation using the Tensorflow library.
+Our implementation is evaluated on both a subset of the ImageNet dataset and a secondary dataset, CIFAR-10, to assess its generalization capability.
+Through this replication and exploration, we seek to validate the core contributions of the Inception model, gain a deeper undertanding of convolutional neural networks, and an appreciation of the associated challenges.
 
@@
 This was one of the key innovations that made GoogLeNet both efficient and powerful, despite being deeper than previous CNN architectures such as AlexNet and VGGNet~\cite{krizhevsky2012imagenet, simonyan2015vgg}.
+
+\subsubsection{Inception Architecture}
+The Inception architecture was introduced to address two primary challenges in CNNs: improving model accuracy while maintaining computational efficiency.
+Szegedy et al.
+proposed a novel structure called the Inception module, which incorporates parallel convolutional filters of different sizes (1x1, 3x3, and 5x5) along with max-pooling.
+The key advantage of this approach is that it allows the network to capture features at multiple scales simultaneously, enhancing the model's ability to learn complex spatial hierarchies within images.
+
+Another major innovation was the use of 1x1 convolutions for dimensionality reduction.
+By applying 1x1 filters before 3x3 and 5x5 convolutions, the Inception module significantly reduces the number of parameters and computational cost without sacrificing accuracy.
+This strategy enables the network to be deeper and wider without a substantial increase in computational complexity.
+
+The Inception architecture was implemented in a 22-layer deep network, GoogLeNet, which was used to win the ILSVRC 2014 competition.
+As Figure \ref{fig:googlenet-comp-perf} shows, GoogLeNet has the lowest top-5 error rate compared to previous winners~\cite{szegedy2015going}.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale=0.7]{project/paper_images/googlenet_competition_perf.png}
+    \caption{GoogLeNet performance at ILSVRC 2014}
+    \label{fig:googlenet-comp-perf}
+\end{figure}
 
@@
 
-After scaling, the images were organized into training, testing, and validation sets. The percentages were 70\% for training, 20\% for testing, and 10\% for validation.
+After scaling, 100 images from each class were set aside to use for evaluating the trained model.
 
-\subsection{A Simple Sequential Network as a Control}
-Before constructing a complex model like GoogLeNet, it is a good idea to use a simple network as a baseline. This network is shown in Figure \ref{fig:cnn-control}
+Training and validation sets were created in-situ using Tensorflow's \textt{image\_dataset\_from\_directory} function. The percentages were 80\% for training, and 20\% for validation.
+
+The networks mentioned in this report were training on an NVIDIA Quadro P2000 GPU having 5GB of dedicated memory.
+
+\subsection{A Simplified GoogLeNet Network}
+Before constructing a complex model like GoogLeNet, it is a good idea to use a simple network as a baseline.
+This network is shown in Figure \ref{fig:cnn-simple}.
+This network is conceptually similar to GoogLeNet but without the complexity of the Inception modules.
+At it's core, the GoogLeNet architecture stacks multiple convolutional layers before downsampling through pooling.
+Stacking multiple layers helps networks capture progressively more complex features, improves generalization, and increases performance while keeping computational cost from getting out of hand.
+This is a common design pattern in many successful image classification networks\cite{simonyan2014very}.
+
+The architecture of this network was informed by the following constraints:
+\begin{itemize}
+    \item 10 image classes versus 1000 image classes in the complete ImageNet dataset.
+    \item About 14.7K images instead of the over 1 million images in the complete ImageNet dataset.
+\end{itemize}
+
+These constraints were the reason why the filter depths range from 32 to 256.
 
@@
     \centering
-    \includegraphics[scale=0.7]{project/paper_images/control_cnn.png}
-    \caption{Simple CNN used as a control}
-    \label{fig:cnn-control}
+    \includegraphics[scale=0.7]{project/paper_images/googlenet_simple.png}
+    \caption{Simplified version of the GoogLeNet network}
+    \label{fig:cnn-simple}
 \end{figure}
 
-This simple model, when trained wfor 25 epochs, achieved a maximum accuracy of 59\% as shown in Figure \ref{fig:cnn-control-accuracy}.
+This simple model, when trained for 30 epochs, achieved a maximum accuracy of 55.8\% at Epoch 29 as shown in Figure \ref{fig:googlenet-simple-accuracy}.
+
+The validation accuracy and training accuracy track well, indicating a lack of overfitting.
+Applying data augmentation to the training set mitigates overfitting, especially with the smaller dataset.
+The augmentation code is shown in Listing \ref{lst:augmentation}
+
+\begin{lstlisting}[caption={Augmentation code}, label={lst:augmentation}]
+augmentation = tf.keras.Sequential([
+    layers.RandomFlip("horizontal"),
+    layers.RandomRotation(0.1),
+    layers.RandomZoom(0.2),
+])
+
+# Apply the augmentation only on the training dataset
+train_dataset = train_dataset.map(lambda x, y: (augmentation(x, training=True), y))
+\end{lstlisting}
 
@@
     \centering
-    \includegraphics[scale=0.7]{project/paper_images/control_cnn_accuracy.png}
-    \caption{Control CNN accuracy}
-    \label{fig:cnn-control-accuracy}
+    \includegraphics[scale=0.7]{project/paper_images/googlenet_simple_accuracy.png}
+    \caption{Accuracy of the simplified version of the GoogLeNet network}
+    \label{fig:googlenet-simple-accuracy}
 \end{figure}
 
-\clearpage
+\subsection{Replicating the GoogLeNet Network}
+The GoogLeNet architecture is explained in tabular form in Table 1 in the Szegedy paper and is reproduced in Figure \ref{fig:googlenet-arch}.
+The introduction of Inception layers having parallel convolutional filters is the primary difference from typical sequential networks.
+The replication attempt, when trained for 30 epochs, achieved a maximum accuracy of 56.9\%.
+This is essentialy the same as the simple network.
+The upshot is that it only took 8 epochs to arrive at the maximum accuracy as shown in Figure \ref{fig:googlenet-complex-accuracy}.
+Nevertheless, both models are nowhere near the accuracy of the model in the Szegedy paper.
+Their model achieved a top-5 error rate of only 6.7\%.
+While the paper does not mention the accuracy, it is not a leap to assume that the accuracy was well above the accuracy in the replication attempt.
+The network is graphically depicted in Figure \ref{fig:googlenet-complex}
 
-\subsection{Replicating the Inception v1 (GoogLeNet) Model}
-The Inception v1 model incorporates branching and parallel operations. As a result, Tensorflow's Sequential  model is inappropriate. Tensorflow's functional API must be used instead.
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale=0.7]{project/paper_images/googlenet_arch.png}
+    \caption{GoogLeNet architecture with Inception layers}
+    \label{fig:googlenet-arch}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale=0.7]{project/paper_images/googlenet_complex_accuracy.png}
+    \caption{Accuracy of the simplified version of the GoogLeNet network}
+    \label{fig:googlenet-complex-accuracy}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{project/paper_images/googlenet_complex.png}
+    \caption{Replicated GoogLeNet network}
+    \label{fig:googlenet-complex}
+\end{figure}
+
+\subsection{Model Evaluation}
+Both networks were evaluated using two sets of images.
+The first is the 1,000 ImageNet images set aside prior to model training.
+The second set is sourced from the CIFAR-10 dataset.
+The CIFAR-10 dataset contains 1,000 32x32 RGB images.
+
+\subsubsection{Performance On Unseen Imagenet Images}
+
+\subsubsection{Performance On CIFAR-10 Images}
+
+\subsection{Challenges}
+The Inception architecture's key contribution is its ability to balance computational efficiency with accuracy.
+By incorporating 1x1 convolutions for dimensionality reduction and parallel convolutions in the Inception modules, GoogLeNet achieves leading performance without increased compute cost or model parameters.
+
+In our reproduction, the model's performance on the ImageNet dataset was not  close to the results reported in the original paper.
+This is not an indictment on the effectiveness of the architecture, however.
+Rather, the small dataset size could be a factor.
+Additionally, our interpretation of the Inception modules to code could have variations from the code used by Szegedy et al.
+
+One of the challenges we encountered was the high computational demand of training, event on the subset of ImageNet images we used.
+Although GoogLeNet is more efficient than many of its contemporaries, training still required GPU resources.
+We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to avoid overfitting.
+We also needed to reduce the learning rate to 0.0001 and reduce batch size to 16 so as not to overwhelm the GPU.
+The training performance is summarized in Table \ref{tab:training_times}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{llll}
+    \hline
+    \textbf{Model} & \textbf{Batch size} & \textbf{Epochs} & \textbf{Training time (minutes)} \\  \hline
+    Simple & 16 & 30 & ?? \\ \hline
+    Complex & 16 & 30 & 69 \\ \hline
+    \end{tabular}
+    \caption{Training times}
+    \label{tab:training_times}
+\end{table}
+
+\subsection{Critical Assessment}
+The Inception architecture has had a significant impact on the field of deep learning, particularly in image classification.
+Its introduction marked a shift from simply increasing the size of models to incorporating more sophisticated designs that optimize the use of computational resources.
+This focus on efficiency has made the architecture suitable for deployment in resource-constrained environments, such as mobile devices.
+
+The architecture has also influenced subsequent innovations, including more advanced object detection systems like Faster R-CNN and SSD, which build on the principles introduced by Inception.
+Additionally, later versions of the Inception architecture continue to push the boundaries of accuracy and efficiency in deep learning.
+
+However, the architecture is not without its limitations.
+Despite its efficient design, training the network still requires significant compute power, especially for large datasets like ImageNet.
+Furthermore, the design of the Inception module, while effective, introduces additional complexity to the network architecture, making it  challenging to implement and optimize.
 

### 2024-09-28 project/utils/file_utils.py
Add script to move random files for evaluation
Created a new utility script in `file_utils.py` to randomly select and move files from the source directory to a target directory. This script helps in preparing evaluation datasets by ensuring a subset of files is moved without manual intervention.

@@ -1,0 +1,20 @@
+import os
+import random
+import shutil
+
+source_dir = os.path.join("../", "data", "imagenet_224")
+target_dir = r"c:\temp\imagenet_224_eval"
+dirs = os.listdir(source_dir)
+num_files = 100
+
+for directory in dirs:
+    files = os.listdir(os.path.join(source_dir, directory))
+    selected_files = random.sample(files, min(num_files, len(files)))
+
+    if not os.path.exists(os.path.join(target_dir, directory)):
+        os.makedirs(os.path.join(target_dir, directory))
+
+    for file in selected_files:
+        source_path = os.path.join(source_dir, directory, file)
+        target_path = os.path.join(target_dir, directory, file)
+        shutil.move(source_path, target_path)

### 2024-09-28 project/paper.tex
Update model evaluation and training details in paper
Revised the accuracy metrics for the simple model and added detailed evaluation results for both ImageNet and CIFAR-10 datasets. Included new references to support dataset and performance analysis, and updated the training configuration to improve GPU performance and resource management.

@@
 
-This simple model, when trained for 30 epochs, achieved a maximum accuracy of 55.8\% at Epoch 29 as shown in Figure \ref{fig:googlenet-simple-accuracy}.
+This simple model, when trained for 30 epochs, achieved a maximum accuracy of 57.2\% at Epoch 23 as shown in Figure \ref{fig:googlenet-simple-accuracy}.
 
@@
 The first is the 1,000 ImageNet images set aside prior to model training.
-The second set is sourced from the CIFAR-10 dataset.
+The second set is sourced from the CIFAR-10 dataset\cite{kim_cifar10_images}.
 The CIFAR-10 dataset contains 1,000 32x32 RGB images.
@@
 \subsubsection{Performance On Unseen Imagenet Images}
+100 images from each class was set aside in order to evaluate the performance of the two models on unseen images. The accuracy of the two models is shown in \ref{tab:perf-unseen-imagenet}.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{ll}
+    \hline
+    \textbf{Model} & \textbf{Accuracy} \\ \hline
+    Simple & 58.1\% \\ \hline
+    GoogLeNet Reproduction & ??? \\ \hline
+    \end{tabular}
+    \caption{Models' performance on unseen ImageNet images}
+    \label{tab:perf-unseen-imagenet}
+\end{table}
+
+The simpler model's accuracy of 58.1\% on the test set is slightly better than the training accuracy of 57.2\%.
+This indicates that the model generalizes well and did not merely memorize the training images.
+In machine learning, when a model performs well on unseen data compared to the training data, it suggests that the model has learned useful patterns that extend beyond the specific examples in the training set.
+
+Furthermore, the relatively small difference in training accuracy and evaluation accuracy indicates that the model has managed to avoid overfitting.
+Overfitting typically results in high training accuracy but significantly lower evaluation accuracy.
+The fact that the model's test accuracy is slightly higher than its training accuracy can be attributed to augmentation done on the training set but not the evaluation set.
+
+On the other hand, the 58.1\% accuracy also indicates that there is room for improvement, suggesting that the model might benefit from more complexity or further tuning.
 
 \subsubsection{Performance On CIFAR-10 Images}
+The CIFAR-10 dataset consists of 1,000 32x32 RGB images (100 images per class). The accuracy of the two models against the unseen CIFAR-10 images is shown in \ref{tab:perf-unseen-imagenet}.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{ll}
+    \hline
+    \textbf{Model} & \textbf{Accuracy} \\ \hline
+    Simple & 21.4\% \\ \hline
+    GoogLeNet Reproduction & ??? \\ \hline
+    \end{tabular}
+    \caption{Models' performance on unseen CIFAR-10 images}
+    \label{tab:perf-unseen-imagenet}
+\end{table}
+
+The low accuracy of 21.4\% achieved by the simple model could be for several reasons.
+First, the models were trained on images having a relatively high resolution compared to CIFAR-10's 32x32 images.
+When models are trained on larger images, the models learn to identify fine-grained details which can become indistinguishable when downscaled to much smaller dimensions~\cite{simonyan2015vgg}.
+
+Additionally, the convolutional filters in the models are likely too large for the small CIFAR-10 images, causing the network to miss out on important features.
+In a 32x32 image, for instance, using large filter sizes or pooling layers can result in excessive downsampling, leading to a significant loss of spatial information, which ultimately hampers the model's ability to classify the images accurately.
+
+Finally, CIFAR-10, like ImageNet, contains complex image classes, which are often difficult to distinguish at such a low resolution.
+The limited information provided by 32x32 images presents additional challenges for models trained on high-resolution images, as they may rely on more detailed textures and edges, which are not as pronounced in low-resolution images~\cite{krizhevsky2012imagenet}.
 
@@
 We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to avoid overfitting.
-We also needed to reduce the learning rate to 0.0001 and reduce batch size to 16 so as not to overwhelm the GPU.
+We also needed to reduce the learning rate to 0.0001, reduced batch size to 16, and enabled memory growth\cite{tensorflow_memory_growth} so as not to overwhelm the GPU.
+Prefetching was also implemented to improve performance\cite{tensorflow_data_performance}.
 The training performance is summarized in Table \ref{tab:training_times}
@@
     \textbf{Model} & \textbf{Batch size} & \textbf{Epochs} & \textbf{Training time (minutes)} \\  \hline
-    Simple & 16 & 30 & ?? \\ \hline
+    Simple & 16 & 30 & 73 \\ \hline
     Complex & 16 & 30 & 69 \\ \hline

### 2024-09-28 project/paper.tex
Update paper and associated references
Corrected typos and clarified details throughout the paper. Modified the methodology and findings sections, improving accuracy statistics and explanations. Updated the bibliography to include a new reference on multi-class image classification using TensorFlow.

@@
 }
-\date{September 20, 2024}
+\date{September 28, 2024}
 
@@
 Our implementation is evaluated on both a subset of the ImageNet dataset and a secondary dataset, CIFAR-10, to assess its generalization capability.
-Through this replication and exploration, we seek to validate the core contributions of the Inception model, gain a deeper undertanding of convolutional neural networks, and an appreciation of the associated challenges.
+Through this replication and exploration, we seek to validate the core contributions of the Inception model, gain a deeper understanding of convolutional neural networks, and an appreciation of the associated challenges.
 
@@
 \subsubsection{Inception Architecture}
-The Inception architecture was introduced to address two primary challenges in CNNs: improving model accuracy while maintaining computational efficiency.
+The Inception architecture was introduced to address the primary challenge in convolutional neural networks- improving model accuracy while maintaining computational efficiency.
 Szegedy et al.
@@
 \subsection{Image Selection}
-The Imagenet dataset contains over 1 million images with a size of over 1TB.
-Processing this amount of data would not be practical given the available coumputing power.
-In order to make processing possible in a reasonable timeframe, a subset of the Imagenet images was downloaded.
+The ImageNet dataset contains over 1 million images with a size of over 1TB.
+Processing this amount of data would not be practical given the available computing power.
+In order to make processing possible in a reasonable time frame, a subset of the ImageNet images was downloaded.
 The subsets, or \textit{synsets} map to the CIFAR-10 image classes as shown in Table \ref{tab:synsets}.
@@
     \end{tabular}
-    \caption{Imagenet synset to CIFAR-10 class mapping}
+    \caption{Mapping ImageNet synsets to CIFAR-10 classes}
     \label{tab:synsets}
@@
 
-The resulting download consists of 14,738 files with a total size of 1.3GB. Scaling the images to 224 x 224 pixels, the dimensions used by Szegedy, et. al. decreased the size to a manageable 137MB.
+The resulting download consists of 14,738 files with a total size of 1.3GB. Scaling the images to 224 x 224 pixels, the dimensions used by Szegedy, et al. decreased the size to a manageable 137MB.
 
@@
 \subsection{A Simplified GoogLeNet Network}
+\label{subsec:simple_model}
 Before constructing a complex model like GoogLeNet, it is a good idea to use a simple network as a baseline.
@@
 This network is conceptually similar to GoogLeNet but without the complexity of the Inception modules.
-At it's core, the GoogLeNet architecture stacks multiple convolutional layers before downsampling through pooling.
+At it's core, the GoogLeNet architecture stacks multiple convolutional layers before down sampling through pooling.
 Stacking multiple layers helps networks capture progressively more complex features, improves generalization, and increases performance while keeping computational cost from getting out of hand.
@@
 
-The validation accuracy and training accuracy track well, indicating a lack of overfitting.
-Applying data augmentation to the training set mitigates overfitting, especially with the smaller dataset.
-The augmentation code is shown in Listing \ref{lst:augmentation}
-
-\begin{lstlisting}[caption={Augmentation code}, label={lst:augmentation}]
-augmentation = tf.keras.Sequential([
-    layers.RandomFlip("horizontal"),
-    layers.RandomRotation(0.1),
-    layers.RandomZoom(0.2),
-])
-
-# Apply the augmentation only on the training dataset
-train_dataset = train_dataset.map(lambda x, y: (augmentation(x, training=True), y))
-\end{lstlisting}
-
+The validation accuracy and training accuracy track well, indicating a lack of overfitting. Applying data augmentation to the training set mitigates overfitting, especially with the smaller dataset.
 \begin{figure}[ht]
@@
 \subsection{Replicating the GoogLeNet Network}
-The GoogLeNet architecture is explained in tabular form in Table 1 in the Szegedy paper and is reproduced in Figure \ref{fig:googlenet-arch}.
+The GoogLeNet architecture is explained in Table 1 of the Szegedy paper is reproduced in Figure \ref{fig:googlenet-arch}.
 The introduction of Inception layers having parallel convolutional filters is the primary difference from typical sequential networks.
-The replication attempt, when trained for 30 epochs, achieved a maximum accuracy of 56.9\%.
-This is essentialy the same as the simple network.
-The upshot is that it only took 8 epochs to arrive at the maximum accuracy as shown in Figure \ref{fig:googlenet-complex-accuracy}.
-Nevertheless, both models are nowhere near the accuracy of the model in the Szegedy paper.
+The replication attempt, when trained for 30 epochs, achieved a maximum accuracy of 60.3\%.
+This is quite improved over the simple model.
+The other impressive aspect is that it only took 14 epochs to arrive at the maximum accuracy as shown in Figure \ref{fig:googlenet-complex-accuracy}.
+The conclusion is that the architecture does provide marked improvements over the simpler, more naive model described in \ref{subsec:simple_model}.
+
+Nevertheless, this model is nowhere near the accuracy of the model in the Szegedy paper.
 Their model achieved a top-5 error rate of only 6.7\%.
 While the paper does not mention the accuracy, it is not a leap to assume that the accuracy was well above the accuracy in the replication attempt.
+We are optimistic, however, that our implementation is sound and that with a larger dataset, the accuracy can be improved and approach the accuracy achieved by Szegedy et.
+al.
 The network is graphically depicted in Figure \ref{fig:googlenet-complex}
@@
     \includegraphics[scale=0.7]{project/paper_images/googlenet_complex_accuracy.png}
-    \caption{Accuracy of the simplified version of the GoogLeNet network}
+    \caption{Accuracy of GoogLeNet network reproduction}
     \label{fig:googlenet-complex-accuracy}
@@
 
-\subsubsection{Performance On Unseen Imagenet Images}
+\subsubsection{Performance On Unseen ImageNet Images}
 100 images from each class was set aside in order to evaluate the performance of the two models on unseen images. The accuracy of the two models is shown in \ref{tab:perf-unseen-imagenet}.
@@
     Simple & 58.1\% \\ \hline
-    GoogLeNet Reproduction & ??? \\ \hline
+    GoogLeNet Reproduction & 66.2\% \\ \hline
     \end{tabular}
@@
 On the other hand, the 58.1\% accuracy also indicates that there is room for improvement, suggesting that the model might benefit from more complexity or further tuning.
+
+The GoogLeNet reproduction resulted in a 66.2\% accuracy on unseen ImageNet images.
+The higher accuracy of this model can be attributed to the same reasons as the simpler model.
+In this case, additional complexity did provide a marked increase in accuracy.
+There is cause for optimism that a result closer to what was achieved by Szegedy, et al.
+could be matched with a larger training set.
 
@@
 First, the models were trained on images having a relatively high resolution compared to CIFAR-10's 32x32 images.
-When models are trained on larger images, the models learn to identify fine-grained details which can become indistinguishable when downscaled to much smaller dimensions~\cite{simonyan2015vgg}.
+When models are trained on larger images, the models learn to identify fine-grained details which can become indistinguishable when down scaled to much smaller dimensions~\cite{simonyan2015vgg}.
 
 Additionally, the convolutional filters in the models are likely too large for the small CIFAR-10 images, causing the network to miss out on important features.
-In a 32x32 image, for instance, using large filter sizes or pooling layers can result in excessive downsampling, leading to a significant loss of spatial information, which ultimately hampers the model's ability to classify the images accurately.
+In a 32x32 image, for instance, using large filter sizes or pooling layers can result in excessive down sampling, leading to a significant loss of spatial information, which ultimately hampers the model's ability to classify the images accurately.
 
-Finally, CIFAR-10, like ImageNet, contains complex image classes, which are often difficult to distinguish at such a low resolution.
+CIFAR-10, like ImageNet, contains complex image classes, which are often difficult to distinguish at such a low resolution.
 The limited information provided by 32x32 images presents additional challenges for models trained on high-resolution images, as they may rely on more detailed textures and edges, which are not as pronounced in low-resolution images~\cite{krizhevsky2012imagenet}.
+
+Finally, Tensorflow up scales the images to 224 x 224 which surely results in a tremendous loss of fidelity.
+The fact that the accuracy is not lower is impressive when considering this loss of fidelity.
 
@@
 Although GoogLeNet is more efficient than many of its contemporaries, training still required GPU resources.
-We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to avoid overfitting.
+We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to avoid overfitting\cite{dishasai_multiclass_classification}.
 We also needed to reduce the learning rate to 0.0001, reduced batch size to 16, and enabled memory growth\cite{tensorflow_memory_growth} so as not to overwhelm the GPU.
 Prefetching was also implemented to improve performance\cite{tensorflow_data_performance}.
-The training performance is summarized in Table \ref{tab:training_times}
+The training performance is summarized in Table \ref{tab:training_times}.
+From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to.
 
@@
     \textbf{Model} & \textbf{Batch size} & \textbf{Epochs} & \textbf{Training time (minutes)} \\  \hline
-    Simple & 16 & 30 & 73 \\ \hline
-    Complex & 16 & 30 & 69 \\ \hline
+    Simple & 16 & 30 & 58 \\ \hline
+    Complex & 16 & 30 & 64 \\ \hline
     \end{tabular}

### 2024-09-29 project/paper.tex
Fix typos and update CIFAR-10 performance table.
Corrected a typo in the TensorFlow dataset function and updated the table reference and accuracy values for CIFAR-10 models. Also adjusted access dates in the bibliography to reflect current information.

@@
 
-Training and validation sets were created in-situ using Tensorflow's \textt{image\_dataset\_from\_directory} function. The percentages were 80\% for training, and 20\% for validation.
+Training and validation sets were created in-situ using Tensorflow's \text{image\_dataset\_from\_directory} function. The percentages were 80\% for training, and 20\% for validation.
 
@@
 \subsubsection{Performance On CIFAR-10 Images}
-The CIFAR-10 dataset consists of 1,000 32x32 RGB images (100 images per class). The accuracy of the two models against the unseen CIFAR-10 images is shown in \ref{tab:perf-unseen-imagenet}.
+The CIFAR-10 dataset consists of 1,000 32x32 RGB images (100 images per class). The accuracy of the two models against the unseen CIFAR-10 images is shown in Table \ref{tab:perf-cifar10}.
 
@@
     Simple & 21.4\% \\ \hline
-    GoogLeNet Reproduction & ??? \\ \hline
+    GoogLeNet Reproduction & 22.6\% \\ \hline
     \end{tabular}
     \caption{Models' performance on unseen CIFAR-10 images}
-    \label{tab:perf-unseen-imagenet}
+    \label{tab:perf-cifar10}
 \end{table}
 
-The low accuracy of 21.4\% achieved by the simple model could be for several reasons.
+The low accuracy of both models could be for several reasons.
 First, the models were trained on images having a relatively high resolution compared to CIFAR-10's 32x32 images.

### 2024-10-10 project/paper.tex
Add Motivation and Inspiration sections to the paper
Updated the paper to include sections on Motivation and Inspiration to give context to the project. Also corrected the date and added new references related to the Inception model implementation and image preprocessing techniques.

@@
 }
-\date{September 28, 2024}
+
+\date{October 11, 2024}
 
@@
 
+\textcolor{red}{
+\section {Motivation}
+Attempting to reproduce the experiment from the Szegedy paper is primarily motivated by the desire to apply TensorFlow to a practical problem, such as image classification, using a realistic dataset.
+Recognizing that the operational environment would be constrained (i.e., running the reproduction on a desktop PC), we also saw this as an opportunity to evaluate the network's performance on a smaller version of the ImageNet dataset.
+This project allowed us to directly experience the challenges of training a neural network.
+Additionally, it provided valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, whereas our prior work has been limited to simpler, sequential networks.}
+
+\textcolor{red}{
+\section{Inspiration}
+Two code bases served as the inspiration for our Tensorflow implementation of Inception.
+The image pre-processing, particularly regularization techniques were guided by \cite{dishasai_multiclass_classification}.
+The Inception layer implementation was inspired by \cite{googlenet_inception_module}.
+Lastly, the graphs in Figures \ref{fig:googlenet-simple-accuracy} and \ref{fig:googlenet-complex-accuracy} were inspired by \cite{datacamp_cnn_tensorflow}.}
+
 \section{Background}
@@
 
-\subsubsection{Inception Architecture}
+\subsection{Inception Architecture}
 The Inception architecture was introduced to address the primary challenge in convolutional neural networks- improving model accuracy while maintaining computational efficiency.

### 2024-10-10 project/paper.tex
Changed red text to black.

@@
 
-\textcolor{red}{
 \section {Motivation}
@@
 This project allowed us to directly experience the challenges of training a neural network.
-Additionally, it provided valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, whereas our prior work has been limited to simpler, sequential networks.}
+Additionally, it provided valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, whereas our prior work has been limited to simpler, sequential networks.
 
-\textcolor{red}{
 \section{Inspiration}
@@
 The Inception layer implementation was inspired by \cite{googlenet_inception_module}.
-Lastly, the graphs in Figures \ref{fig:googlenet-simple-accuracy} and \ref{fig:googlenet-complex-accuracy} were inspired by \cite{datacamp_cnn_tensorflow}.}
+Lastly, the graphs in Figures \ref{fig:googlenet-simple-accuracy} and \ref{fig:googlenet-complex-accuracy} were inspired by \cite{datacamp_cnn_tensorflow}.
 

### 2024-10-10 project/paper.tex
Paper final draft - WIP

--- 2024-10-10 project/paper.tex
@@
 \section {Motivation}
-Attempting to reproduce the experiment from the Szegedy paper is primarily motivated by the desire to apply TensorFlow to a practical problem, such as image classification, using a realistic dataset.
+The primary motivation for attempting to reproduce the experiment from the Szegedy paper was to apply TensorFlow to a practical problem, such as image classification, using a realistic dataset.
+The detailed documentation of the network in the Szegedy paper greatly facilitated its TensorFlow implementation.
 Recognizing that the operational environment would be constrained (i.e., running the reproduction on a desktop PC), we also saw this as an opportunity to evaluate the network's performance on a smaller version of the ImageNet dataset.
-This project allowed us to directly experience the challenges of training a neural network.
-Additionally, it provided valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, whereas our prior work has been limited to simpler, sequential networks.}
+This project provided a hands-on experience with the challenges of training a neural network.
+Additionally, it offered valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, a significant leap from our prior work with simpler sequential networks.
+For comparison, we also experimented with a sequential network to highlight the differences in complexity and performance.}
 
@@
 
-The networks mentioned in this report were training on an NVIDIA Quadro P2000 GPU having 5GB of dedicated memory.
+\textcolor{red}{
+\subsection{Compute Resources}
+The networks created for this project were trained and tested using GPU and CPU resources for comparison purposes.
+The GPU is an NVIDIA Quadro P2000 GPU having 5GB of dedicated memory.
+The CPU is an Intel Core i9-10850K 3.6GHz CPU with 20 logical cores.
+}
 
@@
 
+\textcolor{red}{
+\subsection{Influence on Present Technology}
+Image recognition technology has become ubiquitous and is now readily accessible to the masses.
+For example, Google Photos can organize images based on their content \cite{google_photos_updates_2023}, and one can take a picture of a food label written in a foreign language and translate it using Google Translate \cite{google_translate_support_2024}.
+The success of neural networks can be largely attributed to the influence of Szegedy's work and its experiments.
+However, the original Szegedy experiments required substantial computing power, making them impractical for resource-constrained environments.
+Since 2015, there have been significant advancements in creating deep networks that can efficiently run on memory-limited devices, such as smartphones and automobiles.
+MobileNet, as described in \cite{howard2017mobilenets}, is a class of efficient models designed specifically for mobile applications.
+It builds upon key innovations from Inception, including multi-scale feature extraction, the use of 1x1 convolutions for dimensionality reduction, and the stacking of layers.}
+
 \bibliographystyle{plain}

### 2024-10-10 project/paper.tex
Paper final draft - WIP

@@
 \usepackage{xcolor}
+\usepackage{ulem}
 
@@
 The GoogLeNet architecture is explained in Table 1 of the Szegedy paper is reproduced in Figure \ref{fig:googlenet-arch}.
-The introduction of Inception layers having parallel convolutional filters is the primary difference from typical sequential networks.
+\textcolor{red}{The network is 22 layers deep, incorporating branching and parallel convolutions via Inception modules.} The introduction of Inception layers having parallel convolutional filters is the primary difference from typical sequential networks.
 The replication attempt, when trained for 30 epochs, achieved a maximum accuracy of 60.3\%.
@@
 
-One of the challenges we encountered was the high computational demand of training, event on the subset of ImageNet images we used.
-Although GoogLeNet is more efficient than many of its contemporaries, training still required GPU resources.
+\textcolor{red}{\subsubsection{Computational Demand}}
+\textcolor{red}{One of the challenges we encountered was the high computational demand of training, even on the subset of ImageNet images we used.
+The training performance is summarized in Table \ref{tab:training_times}.
+It is interesting that the training times for the two models were nearly identical when using a \textit{single GPU} but differed by a factor of 1.7 when using \textit{20 logical CPU's.} This drove home one of the reasons GPU's have become so popular.
+From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to.
+To further reduce the time, the learning rate was reduced to 0.0001, and batch size was reduced to 16.
+Prefetching was also implemented to improve performance\cite{tensorflow_data_performance}.} \textcolor{red}{\sout{Although GoogLeNet is more efficient than many of its contemporaries, training still required GPU resources.
 We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to avoid overfitting\cite{dishasai_multiclass_classification}.
@@
 The training performance is summarized in Table \ref{tab:training_times}.
-From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to.
+From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to.}}
 
@@
     \hline
-    \textbf{Model} & \textbf{Batch size} & \textbf{Epochs} & \textbf{Training time (minutes)} \\  \hline
-    Simple & 16 & 30 & 58 \\ \hline
-    Complex & 16 & 30 & 64 \\ \hline
+    \textbf{Model} & \textbf{Batch size} & \textbf{Epochs} & \textbf{Training time (minutes) GPU/CPU} \\  \hline
+    Simple & 16 & 30 & 35/42 \\ \hline
+    GoogLeNet reproduction & 16 & 30 & 37/71 \\ \hline
     \end{tabular}
@@
 \end{table}
+
+\textcolor{red}{\subsubsection{GPU Setup}}
+\textcolor{red}{The setup required to run Tensorflow code on a GPU on Windows is not trivial.
+Fortunately two good resources provided instructions on accomplishing this task \cite{abulhawa_tensorflow_gpu_wsl2_2023} \cite{towardsdatascience_jupyter_wsl2_2023}.}
+
+\textcolor{red}{\subsubsection{Overfitting}}
+\textcolor{red}{Overfitting was also an issue.
+We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to reduce overfitting\cite{dishasai_multiclass_classification}.}
 

### 2024-10-11 project/paper-final.tex
Added final latex and PDF's.

--- 2024-10-10 project/paper.tex
@@
 Additionally, it offered valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, a significant leap from our prior work with simpler sequential networks.
-For comparison, we also experimented with a sequential network to highlight the differences in complexity and performance.}
+For comparison, we also experimented with a sequential network to highlight the differences in complexity and performance.
 
-\textcolor{red}{
 \section{Inspiration}
@@
 
-\textcolor{red}{
 \subsection{Compute Resources}
@@
 The CPU is an Intel Core i9-10850K 3.6GHz CPU with 20 logical cores.
-}
 
@@
 The GoogLeNet architecture is explained in Table 1 of the Szegedy paper is reproduced in Figure \ref{fig:googlenet-arch}.
+The network is 22 layers deep, incorporating branching and parallel convolutions via Inception modules.
 The introduction of Inception layers having parallel convolutional filters is the primary difference from typical sequential networks.
@@
 
-One of the challenges we encountered was the high computational demand of training, event on the subset of ImageNet images we used.
-Although GoogLeNet is more efficient than many of its contemporaries, training still required GPU resources.
-We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to avoid overfitting\cite{dishasai_multiclass_classification}.
-We also needed to reduce the learning rate to 0.0001, reduced batch size to 16, and enabled memory growth\cite{tensorflow_memory_growth} so as not to overwhelm the GPU.
+\subsubsection{Computational Demand}
+One of the challenges we encountered was the high computational demand of training, even on the subset of ImageNet images we used.
+The training performance is summarized in Table \ref{tab:training_times}.
+It is interesting that the training times for the two models were nearly identical when using a \textit{single GPU} but differed by a factor of 1.7 when using \textit{20 logical CPU's.} This drove home one of the reasons GPU's have become so popular.
+From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to.
+To further reduce the time, the learning rate was reduced to 0.0001, and batch size was reduced to 16.
 Prefetching was also implemented to improve performance\cite{tensorflow_data_performance}.
-The training performance is summarized in Table \ref{tab:training_times}.
-From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to.
 
@@
 \end{table}
+
+{\subsubsection{GPU Setup}
+The setup required to run Tensorflow code on a GPU on Windows is not trivial.
+Fortunately two good resources provided instructions on accomplishing this task \cite{abulhawa_tensorflow_gpu_wsl2_2023} \cite{towardsdatascience_jupyter_wsl2_2023}.
+
+\subsubsection{Overfitting}
+Overfitting was also an issue.
+We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to reduce overfitting\cite{dishasai_multiclass_classification}.
 
@@
 
-\textcolor{red}{
 \subsection{Influence on Present Technology}
@@
 MobileNet, as described in \cite{howard2017mobilenets}, is a class of efficient models designed specifically for mobile applications.
-It builds upon key innovations from Inception, including multi-scale feature extraction, the use of 1x1 convolutions for dimensionality reduction, and the stacking of layers.}
+It builds upon key innovations from Inception, including multi-scale feature extraction, the use of 1x1 convolutions for dimensionality reduction, and the stacking of layers.
 

### 2024-10-11 project/paper-final.tex
Add acknowledgments section
Acknowledged Farooq Mahmud for independently completing the project. This includes all tasks related to model implementation, data processing, training, evaluation, and report writing.

@@
 
+\section{Acknowldgements}
+This project was completed solely by\textbf{Farooq Mahmud}.
+All tasks, including the implementation of the TensorFlow code for the GoogLeNet model, data pre-processing, model training and testing, development of the simple sequential model, model evaluation, and writing of the report, were conducted independently.
+
 \bibliographystyle{plain}

### 2024-10-11 project/paper-with-edits.tex
Added final latex and PDF's.
Add acknowledgments section
Acknowledged Farooq Mahmud for independently completing the project. This includes all tasks related to model implementation, data processing, training, evaluation, and report writing.

--- 2024-10-10 project/paper.tex
@@
 
+\textcolor{red}{
+\section{Acknowldgements}
+This project was completed solely by\textbf{Farooq Mahmud}.
+All tasks, including the implementation of the TensorFlow code for the GoogLeNet model, data pre-processing, model training and testing, development of the simple sequential model, model evaluation, and writing of the report, were conducted independently.}
+
 \bibliographystyle{plain}
